---
title: "Integrative GRN inference : permuting the expression and motifs of TFs: effect on expression prediction"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 10)

source('inference_functions/bRF.R')
source('inference_functions/LASSO-D3S.R')
source('inference_functions/evaluateNetwork.R')
source('inference_functions/MSE.R')
# library(ComplexHeatmap)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(ComplexHeatmap)


```

This document demonstrates the use of the bRF and LASSO-D3S functions for integrative GRN inference.

Those functions infer the regulatory pathways of *Arabidopsis thaliana*'s roots in response to nitrate (N) induction from [Varala et al., 2018](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE97500).

They use as inputs the expression profiles of N-responsive genes and TFBS information. 
Prior TFBS information was built by searching in the promoters of the N-responsive genes the PWM of the N-responsive regulators. 

# Data import

Import of the expression data and the N-responsive genes and regulators :

```{r}
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)


ALPHAS <- seq(0,1, by = 0.1)

# subset <- sample(genes, replace = F, size = 20)
subset <- genes

```



# Is the mse variation worsened by unmapping prior values and expression profiles of TFs?


## For Rfs

```{r}
# lmses <- data.frame(row.names = subset)
# N <-100
# for(alpha in ALPHAS){
#   for(perm in 1:N){
#     lmses[,paste(as.character(alpha), perm, "true_data")] <- bRF_inference_MSE(counts, subset, tfs, alpha = alpha, nTrees = 2000,
#                              pwm_occurrence = pwm_occurrence, nCores = 45, tf_expression_permutation = F)
#   }
# 
#   for(perm in 1:N){
#     lmses[,paste(as.character(alpha), perm, "shuffled")] <- bRF_inference_MSE(counts, subset, tfs, alpha = alpha, nTrees = 2000,
#                            pwm_occurrence = pwm_occurrence, nCores = 45, tf_expression_permutation = T)
#   }
# }
# 
# save(lmses, file = "results/brf_perumtations_mse_all_genes.rdata")
<<<<<<< HEAD

=======
load("results/brf_perumtations_mse_all_genes.rdata")
>>>>>>> 9b1478da7002ffb8da85d5b1e86bb4540068c669
# subset<-unique(rownames(lmses))
draw_gene <- function(gene){
  lmses[gene,] %>%
    gather() %>%
    separate(key, into = c("alpha", "rep", "MSEtype"), sep = " ") %>%
    ggplot(aes(x=alpha, y=value, group =interaction(rep, MSEtype), 
               color = MSEtype))+ggtitle(gene)+ylab("MSE/Var(gene)")+
    xlab("alpha")+
    geom_line()+ggtitle(gene)+theme_pubr(legend = "top")
}


draw_gene_mean_sd <- function(gene, title = NULL){
<<<<<<< HEAD
  data <- lmses[gene, ] %>%
=======
  lmses[gene, ] %>%
>>>>>>> 9b1478da7002ffb8da85d5b1e86bb4540068c669
    gather() %>%
    separate(key,
             into = c("alpha", "rep", "MSEtype"),
             sep = " ") %>%
    group_by(alpha, MSEtype) %>%
    mutate(mean_mse = mean(value, na.rm = T),
           sd_mse = sd(value, na.rm = T)) %>%
    ggplot(aes(
      x = as.numeric(alpha),
      y = value,
      color = MSEtype,
      fill = MSEtype
    )) +ggtitle(paste(gene, title))+ylab("MSE/Var(gene)")+
    geom_ribbon(aes(ymin = mean_mse - sd_mse , 
                    ymax = mean_mse + sd_mse  ), 
                alpha = .4)  +theme_pubr(legend = "top")+
    geom_point(alpha = 0.1) + geom_smooth(se=F)+xlab("alpha")

}

get_diff_curves <- function(lmses){
  data <- lmses %>%
  rownames_to_column('gene') %>%
    reshape2::melt()%>%
    separate(variable,
             into = c("alpha", "rep", "MSEtype"),
             sep = " ") %>%
    group_by(gene, alpha, MSEtype) %>%
    mutate(mean_mse = mean(value, na.rm = T),
           sd_mse = sd(value, na.rm = T)) %>%
    dplyr::select(mean_mse, sd_mse, gene, alpha, MSEtype)%>%
    distinct() 
  
  data_true <- filter(data, MSEtype=="true_data")
  data_perm <- filter(data, MSEtype=="shuffled") 
  data_true$mean_mse_perm <- data_perm$mean_mse
  data_true$sd_mse_perm <- data_perm$sd_mse
  return(data_true %>%
    mutate(mean_mse_diff = (mean_mse-mean_mse_perm)/sd_mse_perm))
  # %>%
  #   ggplot(aes(x=as.numeric(alpha), y=mean_mse_diff, color = gene))+
  #   geom_line()
}


# for(gene in sample(genes,50, replace = F)){
#     print(draw_gene(gene)+draw_gene_mean_sd(gene))
# }


```

## For lasso 

```{r, eval=FALSE}
lmses <- data.frame(row.names = subset)
N<-100
for(alpha in ALPHAS){
  # set.seed(121314)
  for(perm in 1:N){
    lmses[,paste(as.character(alpha), perm, "true_data")] <- LASSO.D3S_inference_MSE(counts, subset, tfs, alpha = alpha, N=100,
                             pwm_occurrence = pwm_occurrence, nCores = 40, tf_expression_permutation = F)

    lmses[,paste(as.character(alpha), perm, "shuffled")] <- LASSO.D3S_inference_MSE(counts, subset, tfs, alpha = alpha, N=100,
                           pwm_occurrence = pwm_occurrence, nCores = 40, tf_expression_permutation = T)
  }

}
save(lmses, file = "results/lasso_perumtations_mse_all_genes.rdata")

# load("results/lasso_perumtations_mse_all_genes.rdata")
# for(gene in subset){
#     print(draw_gene(gene) +draw_gene_mean_sd(gene))
# }
```






# Clustering genes for RFs

Based on the difference curves between true and permuted data


```{r}
load("results/brf_perumtations_mse_all_genes.rdata")


diffs <- get_diff_curves(lmses)
diffs %>%
  ggplot(aes(x=as.numeric(alpha), y=mean_mse_diff, group = gene))+
  geom_line(alpha = 0.2)


fractions_out <- diffs %>%
  mutate(diff_greater_than_sd = ifelse(abs(mean_mse_diff)>1, 1, 0)) %>%
  group_by(gene) %>%
  summarise(fraction_out = sum(diff_greater_than_sd)/11);fractions_out<-
  setNames(fractions_out$fraction_out, fractions_out$gene)

diff_curves <- diffs[c("gene", "alpha", "mean_mse_diff")] %>%
  spread(alpha, mean_mse_diff) %>%
  column_to_rownames("gene") %>%
  as.matrix()


diff_curves<-diff_curves[fractions_out[rownames(diff_curves)] > 0,]
  
cor_clust = function(x) hclust(as.dist(1-cor(t(x))), method = "average")

Heatmap(diff_curves, cluster_rows = cor_clust,
        cluster_columns = F, show_row_names = F)


clusters_rf <- cutree(cor_clust(diff_curves), k = 2)
table(clusters_rf)

clusters_rf<- c(clusters_rf,setNames(rep("no diff", sum(fractions_out==0)),
                                     names(fractions_out[fractions_out==0])))
table(clusters_rf)
save(clusters_rf, file = "results/clusters_mse_bRF_100permutations.rdata")

for(gene in sample(genes,40, replace = F)){
    print(draw_gene(gene)+
            draw_gene_mean_sd(gene, title = paste(clusters_rf[gene], round(fractions_out[gene], 4))))
}


for(gene in c("AT1G08090", "AT1G12110")){
    print(draw_gene(gene)+
            draw_gene_mean_sd(gene, title = paste(clusters_rf[gene], round(fractions_out[gene], 4))))
}



ha = HeatmapAnnotation(
    alpha = anno_simple(as.numeric(rep(colnames(diff_curves),1))),
    annotation_name_side = "left")
# draw a heatmap of the genes mean_mse on real data
true_mse <- diffs[c("gene", "alpha", "mean_mse")] %>%
  spread(alpha, mean_mse) %>%
  column_to_rownames("gene") %>%
  as.matrix() 
  

true_mse_pos <- true_mse[names(clusters_rf[clusters_rf==2]),]
Heatmap((true_mse-rowMeans(true_mse))/matrixStats::rowSds(true_mse),
        cluster_columns = F, show_row_names = F, top_annotation = ha)+ 
  rowAnnotation(
    clusters_rf = clusters_rf[rownames(true_mse)],
    col=list(clusters_rf= setNames(c("darkorange", "darkgreen", "lightgrey"), 
                                       nm = names(table(clusters_rf)))))

Heatmap((true_mse_pos-rowMeans(true_mse_pos))/matrixStats::rowSds(true_mse_pos),
        cluster_columns = F, show_row_names = F, top_annotation = ha)+ 
  rowAnnotation(
    clusters_rf = clusters_rf[rownames(true_mse_pos)],
    col=list(clusters_rf= setNames(c("darkorange", "darkgreen", "lightgrey"), 
                                       nm = names(table(clusters_rf)))))
```



# Clustering genes for LASSO

Based on the difference curves between true and permuted data


```{r}
load("results/lasso_perumtations_mse_all_genes.rdata")
diffs <- get_diff_curves(lmses)
diffs %>%
  ggplot(aes(x=as.numeric(alpha), y=mean_mse_diff, group = gene))+
  geom_line(alpha = 0.2)


fractions_out <- diffs %>%
  mutate(diff_greater_than_sd = ifelse(abs(mean_mse_diff)>1, 1, 0)) %>%
  group_by(gene) %>%
  summarise(fraction_out = sum(diff_greater_than_sd)/11);fractions_out<-
  setNames(fractions_out$fraction_out, fractions_out$gene)

diff_curves <- diffs[c("gene", "alpha", "mean_mse_diff")] %>%
  spread(alpha, mean_mse_diff) %>%
  column_to_rownames("gene") %>%
  as.matrix()


diff_curves<-diff_curves[fractions_out[rownames(diff_curves)] > 0,]
  
cor_clust = function(x) hclust(as.dist(1-cor(t(x))), method = "average")

Heatmap(diff_curves, cluster_rows = cor_clust,
        cluster_columns = F, show_row_names = F)


clusters_lasso <- cutree(cor_clust(diff_curves), k = 2)
table(clusters_lasso)

clusters_lasso<- c(clusters_lasso,setNames(rep("no diff", sum(fractions_out==0)),
                                     names(fractions_out[fractions_out==0])))
table(clusters_lasso)
save(clusters_lasso, file = "results/clusters_mse_lasso_100permutations.rdata")

for(gene in sample(genes,20, replace = F)){
    print(draw_gene(gene)+
            draw_gene_mean_sd(gene, title = paste(clusters_lasso[gene], round(fractions_out[gene], 4))))
}

ha = HeatmapAnnotation(
    alpha = anno_simple(as.numeric(rep(colnames(diff_curves),1))),
    annotation_name_side = "left")
# draw a heatmap of the genes mean_mse on real data
true_mse <- diffs[c("gene", "alpha", "mean_mse")] %>%
  spread(alpha, mean_mse) %>%
  column_to_rownames("gene") %>%
  as.matrix() 
  

true_mse_pos <- true_mse[names(clusters_lasso[clusters_lasso==1]),]
Heatmap((true_mse-rowMeans(true_mse))/matrixStats::rowSds(true_mse),
        cluster_columns = F, show_row_names = F, top_annotation = ha)+ 
  rowAnnotation(
    clusters_lasso = clusters_lasso[rownames(true_mse)],
    col=list(clusters_lasso= setNames(c( "darkgreen","darkorange", "lightgrey"), 
                                       nm = names(table(clusters_lasso)))))

Heatmap((true_mse_pos-rowMeans(true_mse_pos))/matrixStats::rowSds(true_mse_pos),
        cluster_columns = F, show_row_names = F, top_annotation = ha)+ 
  rowAnnotation(
    clusters_lasso = clusters_lasso[rownames(true_mse_pos)],
    col=list(clusters_lasso= setNames(c("darkgreen","darkorange", "lightgrey"), 
                                       nm = names(table(clusters_lasso)))))
```

# Functional study of those groups

```{r}
load("rdata/pwm_prom_jaspar_dap.rdata")
load("rdata/gene_structure.rdata")

mean_expr <- rowMeans(counts)[genes]
var_expr <- matrixStats::rowSds(counts[genes,])*matrixStats::rowSds(counts[genes,])
pwm_prom_n_TFs <- pwm_prom[pwm_prom$TF %in% tfs,]

library(patchwork)

# to comment for new version where mse is already normalized per genes
# norm_mse <- exp(as.matrix(cbind(lmses, lmses_lasso)))/var_expr
genes_info <- data.frame(genes = genes, 
                         cluster_rf = clusters_rf[genes],
                         cluster_lasso = clusters_lasso[genes])

genes_info$is_tf <- genes %in% tfs
genes_info$var <- var_expr
genes_info$expr <- mean_expr
genes_info$min_mse <- matrixStats::rowMins(as.matrix(true_mse))
genes_info$nb_motifs <- table(pwm_prom$target)[genes]
genes_info$nb_motifs_n_tfs <- table(pwm_prom_n_TFs$target)[genes]

genes_info[,c("n_introns", "n_transcripts")] <- 
  gene_structure[match(genes_info$gene, gene_structure$gene),
                 c("n_introns", "n_transcripts")]

genes_info%>%
  ggplot(aes(x=cluster_rf, y=log(n_introns))) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Number of introns for RF groups")) +
genes_info%>%
  ggplot(aes(x=cluster_rf, y=n_transcripts)) + 
  geom_boxplot(width=0.1, fill = "white")+
  geom_violin(fill="darkblue", alpha=0.2)+
  ggtitle(("Number of transcripts for RF groups")) + 
  stat_compare_means()

genes_info%>%
  ggplot(aes(x=cluster_rf, y=nb_motifs)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Number of motifs in promoter for RF groups")) + 
  genes_info%>%
  ggplot(aes(x=cluster_rf, y=nb_motifs_n_tfs)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Number of motifs of N-responsive TFs in promoter for RF groups")) + 
  stat_compare_means() + genes_info%>%
  ggplot(aes(x=cluster_rf, y=min_mse)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Min mse for RF groups")) + 
  stat_compare_means() 


genes_info%>%
  ggplot(aes(x=cluster_rf, y=var)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Gene variance for RF groups")) + scale_y_log10()+
  stat_compare_means()+ genes_info%>%
  ggplot(aes(x=cluster_rf, y=expr)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Gene expression for RF groups")) + scale_y_log10()+
  stat_compare_means()


genes_info %>%
  group_by(cluster_rf) %>%
  summarise(n=n(), 
            tf_frac=sum(is_tf)/n()) %>%
  ggplot(aes(x=cluster_rf, y=tf_frac, 
                          label = paste("n=",n))) + 
  geom_bar(stat = "identity", aes(fill=tf_frac), alpha = 1)+
  geom_hline(yintercept = length(tfs)/length(genes)) + 
  geom_text(y=0.2) + xlab("cluster RF") +
  ggtitle("Fraction of TFs in RF groups")+ylim(c(0,0.2))







genes_info%>%
  ggplot(aes(x=cluster_lasso, y=log(n_introns))) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Number of introns for LASSO groups")) +
genes_info%>%
  ggplot(aes(x=cluster_lasso, y=n_transcripts)) + 
  geom_boxplot(width=0.1, fill = "white")+
  geom_violin(fill="darkblue", alpha=0.2)+
  ggtitle(("Number of transcripts for LASSO groups")) + 
  stat_compare_means()

genes_info%>%
  ggplot(aes(x=cluster_lasso, y=nb_motifs)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Number of motifs in promoter for LASSO groups")) + 
  genes_info%>%
  ggplot(aes(x=cluster_lasso, y=nb_motifs_n_tfs)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Number of motifs of N-responsive TFs in promoter for LASSO groups")) + 
  stat_compare_means() + genes_info%>%
  ggplot(aes(x=cluster_lasso, y=min_mse)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Min mse for LASSO groups")) + 
  stat_compare_means() 


genes_info%>%
  ggplot(aes(x=cluster_lasso, y=var)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Gene variance for LASSO groups")) + scale_y_log10()+
  stat_compare_means()+ genes_info%>%
  ggplot(aes(x=cluster_lasso, y=expr)) + 
  geom_violin(fill="darkblue", alpha=0.2)+geom_jitter(width=0.1)+
  geom_boxplot(width=0.1, fill = "white")+
  ggtitle(("Gene expression for LASSO groups")) + scale_y_log10()+
  stat_compare_means()


genes_info %>%
  group_by(cluster_lasso) %>%
  summarise(n=n(), 
            tf_frac=sum(is_tf)/n()) %>%
  ggplot(aes(x=cluster_lasso, y=tf_frac, 
                          label = paste("n=",n))) + 
  geom_bar(stat = "identity", aes(fill=tf_frac), alpha = 1)+
  geom_hline(yintercept = length(tfs)/length(genes)) + 
  geom_text(y=0.2) + xlab("cluster RF") +
  ggtitle("Fraction of TFs in RF groups")+ylim(c(0,0.2))

```




```{r}
# promoteurs enrichis en certains motifs de TFs?
known_tfs <- tfs[which(tfs %in% pwm_prom$TF)]

get_number_of_motifs_per_tfs <- function(genes){
  table(pwm_prom[pwm_prom$target %in% genes & pwm_prom$TF %in% tfs,"TF"])[known_tfs]
}

targets_per_pwm <- data.frame(row.names = known_tfs)
for(group in unique(clusters_rf)){
  targets_per_pwm[paste("lasso", group)] <- get_number_of_motifs_per_tfs(names(
    which(clusters_lasso == group)))/sum(clusters_lasso == group)
  targets_per_pwm[paste("rf", group)] <- get_number_of_motifs_per_tfs(names(
    which(clusters_rf == group)))/sum(clusters_rf == group)

}

enrichments_per_pwm <- data.frame(row.names = known_tfs)
n_genes <- length(genes)
for(group in unique(clusters_rf)){
  # number of motifs in all the genes
  n_targets_lasso_in_all <- get_number_of_motifs_per_tfs(genes)
  
  n_targets_lasso_in_group <- get_number_of_motifs_per_tfs(names(
    which(clusters_lasso == group)))
  n_group_lasso <- length(names(which(clusters_lasso == group)))
  
  n_targets_rf_in_group <- get_number_of_motifs_per_tfs(names(
    which(clusters_rf == group)))
  n_group_rf <- length(names(which(clusters_rf == group)))
  
  for(tf in known_tfs){
    
    # number of genes with that motif in all genes
    n_targets_in_all_tf <- n_targets_lasso_in_all[tf]
    
    # number of genes with that motif in the lasso group
    n_targets_lasso_in_group_tf <- n_targets_lasso_in_group[tf]
    p_lasso <- phyper(q=n_targets_lasso_in_group_tf-1,
           m=n_targets_in_all_tf, #white balls
           n=n_genes-n_targets_in_all_tf, # black balls
           k=n_group_lasso, lower.tail = FALSE)
    
     # number of genes with that motif in the rf group
    n_targets_rf_in_group_tf <- n_targets_rf_in_group[tf]
    p_rf <- phyper(q=n_targets_rf_in_group_tf-1,
           m=n_targets_in_all_tf, 
           n=n_genes-n_targets_in_all_tf,
           k=n_group_rf, lower.tail = FALSE)
    
    enrichments_per_pwm[tf, paste0(group, "lasso")]<- p_lasso
    enrichments_per_pwm[tf, paste0(group, "rf")]<- p_rf
  }
}

enrichments_per_pwm[enrichments_per_pwm<0.05] <- 0
enrichments_per_pwm[enrichments_per_pwm>=0.05] <- 1
Heatmap(enrichments_per_pwm, cluster_columns = F)


tfs_rf_pwm_pos <- rownames(enrichments_per_pwm[enrichments_per_pwm$`2rf`==0,])
tfs_lasso_pwm_pos <- rownames(enrichments_per_pwm[enrichments_per_pwm$`1lasso`==0,])
DIANE::get_gene_information(tfs_rf_pwm_pos, organism = "Arabidopsis thaliana")
DIANE::get_gene_information(tfs_lasso_pwm_pos, organism = "Arabidopsis thaliana")
DIANE::get_gene_information(intersect(tfs_rf_pwm_pos, tfs_lasso_pwm_pos ), organism = "Arabidopsis thaliana")

tfs_rf_pwm_bad <- rownames(enrichments_per_pwm[enrichments_per_pwm$`1rf`==0,])
tfs_lasso_pwm_bad <- rownames(enrichments_per_pwm[enrichments_per_pwm$`2lasso`==0,])
DIANE::get_gene_information(tfs_rf_pwm_bad, organism = "Arabidopsis thaliana")
DIANE::get_gene_information(tfs_lasso_pwm_bad, organism = "Arabidopsis thaliana")

tfs_rf_pwm_pretty_bad <- rownames(enrichments_per_pwm[enrichments_per_pwm$`no diffrf`==0,])
tfs_lasso_pwm_pretty_bad <- rownames(enrichments_per_pwm[enrichments_per_pwm$`no difflasso`==0,])
DIANE::get_gene_information(tfs_rf_pwm_pretty_bad, organism = "Arabidopsis thaliana")
DIANE::get_gene_information(tfs_lasso_pwm_pretty_bad, organism = "Arabidopsis thaliana")
```


# Network ranks of the Tfs that have enriched motifs in some gene clusters



```{r, fig.width=10, fig.height=10, eval=FALSE}
load("results/100_permutations_bRF_edges.rdata")
# 10 first replicates
edges <- edges[names(edges)[as.numeric(str_split_fixed(names(edges), '_', 5)[,4])<=10]]
net <- edges$bRF_1_trueData_1_0.005
# nodes <- 
get_nodes_relative_rank <- function(net, nodes){
  return(setNames(rank(table(net$from))[nodes]/length(unique(net$from)),nodes))
}

positive_genes <- names(clusters_rf[clusters_rf==2])
edges_positives <- lapply(edges, function(net){net[net$to %in% positive_genes,]})

negative_genes <- names(clusters_rf[clusters_rf==1])
edges_negatives <- lapply(edges, function(net){net[net$to %in% negative_genes,]})


relative_rank <- lapply(edges, get_nodes_relative_rank, tfs_rf_pwm_pos)
relative_rank_pos <- lapply(edges_positives, get_nodes_relative_rank, tfs_rf_pwm_pos)

data.frame(relative_rank) %>%
  rownames_to_column("gene") %>%
  reshape2::melt() %>%
  separate(variable, into = c("method", "alpha", "dataset", "rep", "density"), sep = '_', remove = F) %>%
  mutate(alpha = as.numeric(alpha)) %>%
  ggplot(aes(x=alpha, y=value, color = dataset)) + 
  ggh4x::facet_nested_wrap(vars(density, gene), nest_line = T) + 
  geom_point() + geom_smooth() +
  theme(strip.background = element_blank(), axis.title.x = element_text(size = 22),
        title = element_text(size = 16), strip.text = element_text(size = 16), 
        legend.text = element_text(size = 15), axis.text = element_text(size = 12), 
        legend.position = 'left') + ggtitle("TFs with motifs enriched in positive genes promoters : relative out-degree rank")


data.frame(relative_rank_pos) %>%
  rownames_to_column("gene") %>%
  reshape2::melt() %>%
  separate(variable, into = c("method", "alpha", "dataset", "rep", "density"), sep = '_', remove = F) %>%
  mutate(alpha = as.numeric(alpha)) %>%
  ggplot(aes(x=alpha, y=value, color = dataset)) + 
  ggh4x::facet_nested_wrap(vars(density, gene), nest_line = T) + 
  geom_point() + geom_smooth() +
  theme(strip.background = element_blank(), axis.title.x = element_text(size = 22),
        title = element_text(size = 16), strip.text = element_text(size = 16), 
        legend.text = element_text(size = 15), axis.text = element_text(size = 12), 
        legend.position = 'left') + ggtitle("TFs with motifs enriched in positive genes promoters : relative out-degree rank in positive subset")



relative_rank_bad <- lapply(edges, get_nodes_relative_rank, tfs_rf_pwm_bad)
relative_rank_bad_subset <- lapply(edges_negatives, get_nodes_relative_rank, tfs_rf_pwm_bad)

data.frame(relative_rank_bad) %>%
  rownames_to_column("gene") %>%
  reshape2::melt() %>%
  separate(variable, into = c("method", "alpha", "dataset", "rep", "density"), sep = '_', remove = F) %>%
  mutate(alpha = as.numeric(alpha)) %>%
  ggplot(aes(x=alpha, y=value, color = dataset)) + 
  ggh4x::facet_nested_wrap(vars(density, gene), nest_line = T) + 
  geom_point() + geom_smooth() +
  theme(strip.background = element_blank(), axis.title.x = element_text(size = 22),
        title = element_text(size = 16), strip.text = element_text(size = 16), 
        legend.text = element_text(size = 15), axis.text = element_text(size = 12), 
        legend.position = 'left') + ggtitle("TFs with motifs enriched in negative genes promoters : relative out-degree rank")

data.frame(relative_rank_bad_subset) %>%
  rownames_to_column("gene") %>%
  reshape2::melt() %>%
  separate(variable, into = c("method", "alpha", "dataset", "rep", "density"), sep = '_', remove = F) %>%
  mutate(alpha = as.numeric(alpha)) %>%
  ggplot(aes(x=alpha, y=value, color = dataset)) + 
  ggh4x::facet_nested_wrap(vars(density, gene), nest_line = T) + 
  geom_point() + geom_smooth() +
  theme(strip.background = element_blank(), axis.title.x = element_text(size = 22),
        title = element_text(size = 16), strip.text = element_text(size = 16), 
        legend.text = element_text(size = 15), axis.text = element_text(size = 12), 
        legend.position = 'left') + ggtitle("TFs with motifs enriched in negative genes promoters : relative out-degree rank in negative subset")

```



# Go enrichments

```{r, eval=T}
library(DIANE)
background <- convert_from_agi(genes)

for(group in unique(clusters_rf)){
  
  genes_i <- names(which(clusters_lasso == group))

  print(paste("LASSO", length(genes_i), "genes,", group, "\n"))
  genes_i <- convert_from_agi(genes_i)
  go_lasso <- enrich_go(genes_i, background)
  DIANE::draw_enrich_go(go_lasso)
  print(go_lasso)
  genes_i <- names(which(clusters_rf == group))
  print(paste("RF", length(genes_i), "genes, group", group))
  genes_i <- convert_from_agi(genes_i)
  go_rf <- enrich_go(genes_i, background)
  DIANE::draw_enrich_go(go_rf)
  print(go_rf)
}



```

# Expression


```{r}
expression <- data.frame(counts)
expression <- (expression-rowMeans(expression)) / matrixStats::rowSds(as.matrix(expression))

Heatmap(expression, cluster_columns = F, show_row_names = F)+
  rowAnnotation(
    clusters_rf = clusters_rf[rownames(expression)],
    clusters_lasso = clusters_lasso[rownames(true_mse)],
    col=list(clusters_lasso= setNames(c( "darkgreen","darkorange", "lightgrey"), 
                                       nm = names(table(clusters_lasso))),
      clusters_rf= setNames(c("darkorange", "darkgreen", "lightgrey"), 
                                       nm = names(table(clusters_rf)))))

```


# intersection of groups

```{r}
load("results/clusters_mse_bRF_100permutations.rdata")
load("results/clusters_mse_lasso_100permutations.rdata")
DIANE::draw_venn(list("positive for lasso" = names(clusters_lasso[clusters_lasso==1]),
                      "positive for RFs" = names(clusters_rf[clusters_rf==2])))

DIANE::draw_venn(list("negative for lasso" = names(clusters_lasso[clusters_lasso==2]),
                      "negative for RFs" = names(clusters_rf[clusters_rf==1])))

DIANE::draw_venn(list("no diff for lasso" = names(clusters_lasso[clusters_lasso=="no diff"]),
                      "no diff for RFs" = names(clusters_rf[clusters_rf=="no diff"])))
```






