---
title: "Integrative GRN inference tutorial"
output: html_document
---

Before being able to use bRF, the modified C++ Random Forest implementation have to be set up on your computer. To do so, run in your system terminal :

```R CMD SHLIB inference_functions/Cpp_dependencies/rfutils.c```
```R CMD SHLIB inference_functions/Cpp_dependencies/regTree.c```
```R CMD SHLIB inference_functions/Cpp_dependencies/regrf_mse.c```

The R packages that you will need to install are :
doParallel, parallel, foreach, doRNG, randomForest, igraph, boot, stringr, tictoc, GENIE3

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source('inference_functions/bRF.R')
source('inference_functions/LASSO-D3S.R')
source('inference_functions/evaluateNetwork.R')
source('inference_functions/MSE.R')
```

This document demonstrates the use of the bRF and LASSO-D3S functions for integrative GRN inference.

Those functions infer the regulatory pathways of *Arabidopsis thaliana*'s roots in response to nitrate (N) induction from [Varala et al., 2018](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE97500).

They use as inputs the expression profiles of N-responsive genes and TFBS information. 
Prior TFBS information was built by searching in the promoters of the N-responsive genes the PWM of the N-responsive regulators. 

## Data import

### Expression data

Import of the expression data and the N-responsive genes and regulators :

```{r}
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
```

### TFBS data


```{r}
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)
```

## GRN inference


### bRf : biased Random Forests

Let's infer a GRN using bRF with $\alpha=0.8$ and $D = 0.005$.

```{r}
importances <- bRF_inference(counts, genes, tfs, alpha = 0.8, 
                             pwm_occurrence = pwm_occurrence, nCores = 7)
grn_bRF <- bRF_network(importances, density = 0.005, pwm_occurrence, genes, tfs)
head(grn_bRF); dim(grn_bRF)
mean(grn_bRF$pwm)
```


### LASSO-D3S : LASSO with Differential Shrinkage and Stability Selection

Let's infer a GRN using LASSO-D3S with $\alpha=0.5$ and $D = 0.005$.

```{r}
pvalues <- LASSO.D3S_inference(counts, genes, tfs, alpha = 0.5, N = 50,
                               pwm_occurrence = pwm_occurrence, nCores = 7)
grn_lasso.d3s <- LASSO.D3S_network(pvalues, density = 0.005, pwm_occurrence, genes, tfs)
head(grn_lasso.d3s); dim(grn_lasso.d3s)
mean(grn_lasso.d3s$pwm)
```


### Validating the GRNs against ConnecTF

```{r}
connecTF_validation <- evaluate_networks(list("bRF" = grn_bRF, "LASSO-D3S" = grn_lasso.d3s), 
                                         input_genes = genes, input_tfs = tfs, 
                                         nCores =2);connecTF_validation
```

In this example, approximately approximately 40% of predicted edges by bRF and LASSO-D3S are supported by ConnecTF experiments.


### Computing prediction error

```{r}
get_MSE(grn_bRF, method = "bRF", nCores=6)
get_MSE(grn_lasso.d3s, method = "LASSO-D3S", nCores=6)
```


