---
title: "Integrative GRN inference : permuting the expression and motifs of TFs: effect on expression prediction"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 10)

source('inference_functions/bRF.R')
source('inference_functions/LASSO-D3S.R')
source('inference_functions/evaluateNetwork.R')
source('inference_functions/MSE.R')
# library(ComplexHeatmap)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(ComplexHeatmap)


```

This document demonstrates the use of the bRF and LASSO-D3S functions for integrative GRN inference.

Those functions infer the regulatory pathways of *Arabidopsis thaliana*'s roots in response to nitrate (N) induction from [Varala et al., 2018](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE97500).

They use as inputs the expression profiles of N-responsive genes and TFBS information. 
Prior TFBS information was built by searching in the promoters of the N-responsive genes the PWM of the N-responsive regulators. 

# Data import

Import of the expression data and the N-responsive genes and regulators :

```{r}
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)


ALPHAS <- seq(0,1, by = 0.1)

# subset <- sample(genes, replace = F, size = 20)
subset <- genes


load("results/brf_perumtations_mse_all_genes.rdata")


load("results/100_permutations_bRF_importances.rdata")
```



# Is the mse variation worsened by unmapping prior values and expression profiles of TFs?


## For Rfs

```{r, message=F}

draw_gene <- function(gene){
  lmses[gene,] %>%
    gather() %>%
    separate(key, into = c("alpha", "rep", "MSEtype"), sep = " ") %>%
    ggplot(aes(x=alpha, y=value, group =interaction(rep, MSEtype), 
               color = MSEtype))+ggtitle(gene)+ylab("MSE/Var(gene)")+
    xlab("alpha")+
    geom_line()+ggtitle(gene)+theme_pubr(legend = "top")
}


draw_gene_mean_sd <- function(gene, title = NULL){
  lmses[gene, ] %>%
    gather() %>%
    separate(key,
             into = c("alpha", "rep", "MSEtype"),
             sep = " ") %>%
    group_by(alpha, MSEtype) %>%
    mutate(mean_mse = mean(value, na.rm = T),
           sd_mse = sd(value, na.rm = T)) %>%
    ggplot(aes(
      x = as.numeric(alpha),
      y = value,
      color = MSEtype,
      fill = MSEtype
    )) +ggtitle(paste(gene, title))+ylab("MSE/Var(gene)")+
    geom_ribbon(aes(ymin = mean_mse - sd_mse , 
                    ymax = mean_mse + sd_mse  ), 
                alpha = .4)  +theme_pubr(legend = "top")+
    geom_point(alpha = 0.1) + geom_smooth(se=F)+xlab("alpha")

}



draw_gene_effective_integration <- function(gene){
  tfs_with_motif <- names(which(pwm_occurrence[gene,]==1))
  data.frame(lapply(mats, function(mat){mean(rank(mat[,gene])[tfs_with_motif])})) %>%
  gather(key = "setting", value = "summed_importance") %>%
  separate(setting, into = c("method", "alpha", "dataset", "rep"), sep = "_") %>%
  group_by(alpha, dataset) %>%
  mutate(mean_imp = mean(summed_importance),
            sd_imp = sd(summed_importance),
         alpha = as.numeric(alpha))%>%
  ggplot(aes(x=alpha, y = summed_importance, color = dataset, fill = dataset)) +
  geom_point(alpha = 0.1) + geom_smooth(se=F)+
  geom_ribbon(aes(ymin = mean_imp - sd_imp , 
                    ymax = mean_imp + sd_imp  ), 
                alpha = .4)  +theme_pubr(legend = "top") + 
    ylab("effective data integration")
}

draw_MSE_vs_gene_effective_integration <- function(gene){
  tfs_with_motif <- names(which(pwm_occurrence[gene,]==1))
  
  inte <- data.frame(lapply(mats, function(mat){mean(rank(mat[,gene])[tfs_with_motif])})) %>%
  gather(key = "setting", value = "summed_importance") %>%
  separate(setting, into = c("method", "alpha", "dataset", "rep"), sep = "_") %>%
  group_by(alpha, dataset) %>%
  mutate(mean_imp = mean(summed_importance),
            sd_imp = sd(summed_importance),
         alpha = as.numeric(alpha))
  
  
  lmses[gene, ] %>%
    gather() %>%
    separate(key,
             into = c("alpha", "rep", "MSEtype"),
             sep = " ") %>%
    group_by(alpha, MSEtype) %>%
    rename(dataset = MSEtype)%>%
    mutate(mean_mse = mean(value, na.rm = T),
           sd_mse = sd(value, na.rm = T),
           alpha = as.numeric(alpha),
           dataset = str_replace(dataset, "true_data", "trueData")) %>%
    full_join(inte, by = c("alpha", "dataset", "rep")) %>%
  ggplot(aes(y=value, x = summed_importance, color = dataset, fill = dataset)) +
  geom_point(alpha = 0.1, size = 0.5)+theme_pubr(legend = "top") + geom_smooth()+
    ylab("MSE/Var(gene)") + xlab("effective integration") 
}

draw_MSE_vs_gene_effective_integration_means <- function(gene){
  tfs_with_motif <- names(which(pwm_occurrence[gene,]==1))
  
  inte <- data.frame(lapply(mats, function(mat){mean(rank(mat[,gene])[tfs_with_motif])})) %>%
  gather(key = "setting", value = "summed_importance") %>%
  separate(setting, into = c("method", "alpha", "dataset", "rep"), sep = "_") %>%
  group_by(alpha, dataset) %>%
  summarise(mean_imp = mean(summed_importance),
            sd_imp = sd(summed_importance),
         alpha = as.numeric(alpha))
  
  
  lmses[gene, ] %>%
    gather() %>%
    separate(key,
             into = c("alpha", "rep", "MSEtype"),
             sep = " ") %>%
    group_by(alpha, MSEtype) %>%
    rename(dataset = MSEtype)%>%
    summarise(mean_mse = mean(value, na.rm = T),
           sd_mse = sd(value, na.rm = T),
           alpha = as.numeric(alpha),
           dataset = str_replace(dataset, "true_data", "trueData")) %>%
    full_join(inte, by = c("alpha", "dataset")) %>%
  ggplot(aes(y=mean_mse, x = mean_imp, group = dataset, color = dataset, fill = dataset)) +
  geom_point(alpha = 0.1)+theme_pubr(legend = "top") +  geom_line()+
    ylab("MSE/Var(gene)") + xlab("effective integration") 
}



for(gene in sample(genes,20, replace = F)){
    print(draw_gene_effective_integration(gene)+draw_gene_mean_sd(gene)+draw_MSE_vs_gene_effective_integration(gene))
}


```


On the class of interest:

```{r}
load("results/clusters_mse_bRF_100permutations.rdata")
pos_genes <- names(which(clusters_rf==2))
for(gene in sample(pos_genes,20, replace = F)){
    print(draw_gene_effective_integration(gene)+draw_gene_mean_sd(gene)+draw_MSE_vs_gene_effective_integration(gene))
}
```

