---
title: "Integrative GRN inference : model exploration"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 2
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

source('inference_functions/bRF.R')
source('inference_functions/LASSO-D3S.R')
source('inference_functions/evaluateNetwork.R')
source('inference_functions/MSE.R')
library(ComplexHeatmap)
library(ggplot2)
library(tidyverse)
library(ggpubr)
```

This document demonstrates the use of the bRF and LASSO-D3S functions for integrative GRN inference.

Those functions infer the regulatory pathways of *Arabidopsis thaliana*'s roots in response to nitrate (N) induction from [Varala et al., 2018](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE97500).

They use as inputs the expression profiles of N-responsive genes and TFBS information. 
Prior TFBS information was built by searching in the promoters of the N-responsive genes the PWM of the N-responsive regulators. 

## Data import

### Expression data and TFBS data

Import of the expression data and the N-responsive genes and regulators :

```{r}
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)
```



# Is the mse variation worsened by unmapping prior values and expression profiles of TFs?


## Exploring the intrinsic variation of MSE profiles per genes, no permutations

Without setting any seed, mse profiles are quite similar.

```{r}
ALPHAS <- seq(0,1, by = 0.1)

subset <- sample(genes, replace = F, size = 20)
lmses <- data.frame(row.names = subset)
lmses_perm<- data.frame(row.names = subset)

for(alpha in ALPHAS){
  # set.seed(121314)
  lmses[,as.character(alpha)] <- bRF_inference_MSE(counts, subset, tfs, alpha = alpha, nTrees = 4000,
                             pwm_occurrence = pwm_occurrence, nCores = 15)
  # set.seed(121314)
  lmses_perm[,as.character(alpha)] <- bRF_inference_MSE(counts, subset, tfs, alpha = alpha, nTrees = 4000,
                           pwm_occurrence = pwm_occurrence, nCores = 15, tf_expression_permutation = F)
  
}


draw_gene <- function(gene){
  df <- data.frame("mse1"=as.numeric(lmses[gene,]), 
                   "mse2"=as.numeric(lmses_perm[gene,]), alpha = ALPHAS)
  gather(df, mse1, mse2, key = "permutation", value = "mse") %>%
    ggplot(aes(x=alpha, y=mse, group =permutation, color = permutation))+
             geom_line()
}
for(gene in subset){
    print(draw_gene(gene))
}

```

```{r}
lmses <- data.frame(row.names = subset)
lmses_perm<- data.frame(row.names = subset)


for(alpha in ALPHAS){
  # set.seed(121314)
  lmses[,as.character(alpha)] <- LASSO.D3S_inference_MSE(counts, subset, tfs, alpha = alpha, N=200,
                             pwm_occurrence = pwm_occurrence, nCores = 15)
  lmses_perm[,as.character(alpha)] <- LASSO.D3S_inference_MSE(counts, subset, tfs, alpha = alpha, N=200,
                           pwm_occurrence = pwm_occurrence, nCores = 15, tf_expression_permutation = F)

}


for(gene in subset){
    print(draw_gene(gene))
}
```

# Comparing real MSE curves with 10 permutated ones


## For Rfs

```{r}
lmses <- data.frame(row.names = subset)
lmses_perm<- data.frame(row.names = subset)

for(alpha in ALPHAS){
  lmses[,as.character(alpha)] <- bRF_inference_MSE(counts, subset, tfs, alpha = alpha, nTrees = 4000,seed =121314,
                             pwm_occurrence = pwm_occurrence, nCores = 15, tf_expression_permutation = F)
  
  for(perm in 1:10){
    lmses_perm[,paste(as.character(alpha), perm)] <- bRF_inference_MSE(counts, subset, tfs, alpha = alpha, nTrees = 4000,
                           pwm_occurrence = pwm_occurrence, nCores = 15, tf_expression_permutation = T, seed = 121314)
  }
}

draw_gene <- function(gene){
  
  cbind.data.frame(lmses[gene,], lmses_perm[gene,]) %>%
    gather() %>%
    separate(key, into = c("alpha", "rep"), sep = " ") %>%
    mutate(permutated = !is.na(rep))%>%
    ggplot(aes(x=alpha, y=value, group =rep, 
               color = permutated))+
    geom_line()
}

# draw_gene <- function(gene){
#   
#   cbind.data.frame(lmses[gene,], lmses_perm[gene,]) %>%
#     gather() %>%
#     separate(key, into = c("alpha", "rep"), sep = " ") %>%
#     mutate(permutated = !is.na(rep))%>%
#     group_by(alpha, permutated) %>%
#     summarise(mean_mse = mean(value, na.rm=T),
#               std_mse = sd(value, na.rm=T))%>%
#     ggplot(aes(x=alpha, y=mean_mse,
#                color = permutated))+
#     geom_point()
# }

for(gene in subset){
    print(draw_gene(gene))
}
```

## For lasso 

```{r}

lmses <- data.frame(row.names = subset)
lmses_perm<- data.frame(row.names = subset)

for(alpha in ALPHAS){
  # set.seed(121314)
  lmses[,as.character(alpha)] <- LASSO.D3S_inference_MSE(counts, subset, tfs, alpha = alpha, N=200,seed=121314,
                             pwm_occurrence = pwm_occurrence, nCores = 15, tf_expression_permutation = F)
  
  for(perm in 1:10){
    lmses_perm[,paste(as.character(alpha), perm)] <- LASSO.D3S_inference_MSE(counts, subset, tfs, alpha = alpha, N=200,seed = 121314,
                           pwm_occurrence = pwm_occurrence, nCores = 15, tf_expression_permutation = T)
  }
}


for(gene in subset){
    print(draw_gene(gene))
}
```





























