---
title: "Comparing DIOgene to other existing implementations for integrative GRN inference"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 10)

source('inference_functions/weightedRF.R')
source('inference_functions/weightedLASSO.R')
source('inference_functions/evaluateNetwork.R')
source('inference_functions/data_integration_optimization.R')

library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(ggVennDiagram)
library(ComplexHeatmap)
library(clusterProfiler)
library(circlize)
```


# Non linear case: comparison to iRafNet

Differences between iRafNet and weightedRF:

+ Importance metric (Node Purity vs permutation approach)
+ iRafNet has to scale the features to get appropriate feature importance


```{r}
# orifinal iRafNet, simply with the MSE having been added as an output
# library(iRafNet)

# building expression
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
# x <- t(counts)[,genes]

# building prior (for three values of alpha: 0, 0.5, 1)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)
# adding the priors for gene-gene interactions only
# targets_prior <- matrix(0,length(genes),length(genes)-length(tfs))
# colnames(targets_prior) <- setdiff(genes, tfs)
# rownames(targets_prior) <- genes

# pwm_imputed <- pwm_occurrence
# pwm_imputed[is.na(pwm_imputed)] <- 0.5
# 
# # link function between priors and subsampling weights
# get_weight <- function(pi, alpha_gene){
#   return(ifelse(pi == 1, sqrt(1-(alpha_gene-1)^2)+1,
#          ifelse(pi == 0.5, 1-alpha_gene,
#                 -sqrt(1-(alpha_gene-1)^2)+1)))
# }
# 
# get_weights <- function(alpha_gene){
#   W_tfs <- apply(pwm_imputed, 1:2, 
#                  FUN=get_weight, alpha_gene=alpha_gene)
#   W <- cbind(targets_prior, W_tfs)[genes,genes]
#   return(W)
# }
# 
# priors <- get_weights(alpha_gene = 0.5)

alphas <- c(0, 0.5, 1)
results <- NULL
for(alpha in alphas){
  # computing models for iRafNet and weightedRF
  importances_iRafNet <- weightedRF_inference(counts, genes, tfs, alpha = alpha, nTrees = 2000,
                                       importance = "IncNodePurity", pwm_occurrence = pwm_occurrence, 
                                       nCores = 10)

  importances_wRF <- weightedRF_inference(counts, genes, tfs, alpha = alpha, nTrees = 2000,
                                         pwm_occurrence = pwm_occurrence, 
                                         nCores = 10)
  
  # building sparse GRNs and evaluating them against DAP-Seq
  pr_curve_iRafNet <- evaluate_fully_connected(importances_iRafNet[tfs,], 
                                               pwm_occurrence = pwm_occurrence,
                            input_genes = genes, input_tfs = tfs, 
                                           validation = c("DAPSeq"),
                                           nCores =10) %>%
    mutate(alpha = alpha, method = "iRafNet", 
           mse = median(importances_iRafNet["mse",], na.rm = T))
  
  pr_curve_wRF <- evaluate_fully_connected(importances_wRF[tfs,], 
                                           pwm_occurrence = pwm_occurrence,
                            input_genes = genes, input_tfs = tfs, 
                                           validation = c("DAPSeq"),
                                           nCores =10) %>%
    mutate(alpha = alpha, method = "weightedRF",
           mse = median(importances_wRF["mse",], na.rm = T))
  
  if(is.null(results)) results <- rbind.data.frame(pr_curve_iRafNet, pr_curve_wRF)
  else results <- rbind(results, rbind.data.frame(pr_curve_iRafNet, pr_curve_wRF))
}

(results %>%
  ggplot(aes(y=precision, x = recall, color = method)) +
  geom_point() + geom_line()+
  ggh4x::facet_nested_wrap(vars(paste(alpha)))+
  theme_bw() )/(
results %>%
  group_by(alpha, method) %>%
  distinct(mse, .keep_all = T) %>%
  ggplot(aes(y=mse, x =method, fill = method)) +
  geom_bar(stat = "identity") + 
  ggh4x::facet_nested_wrap(vars(paste(alpha)))+
  theme_bw())

results %>%
  ggplot(aes(y=pwm_supprt, x =alpha, color = method)) +
  geom_point() + geom_line()+
  ggh4x::facet_nested_wrap(vars(paste(network_name)), ncol = 7)+
  theme_bw()


save(results, file = "results/irafnet_weightedRF_comparison.rdata")
```

# Linear case : comparison to stars-lasso in the Inferlator?

Generating input tsv files to feed the Inferelator worker:

```{r}
# expression file : 
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)

# expression should be log transformed as the model is 
expression <- log10(counts + 0.75) %>%
  as.data.frame() 
options(OutDec=".")
# %>%
#   rownames_to_column("gene")

write.table(expression, file = "data/inferelator_inputs/expression.tsv", 
            sep = '\t', row.names = T, quote = F)

# names of the TFs
write.table(tfs, file = "data/inferelator_inputs/tf_names.tsv", sep = '\t', 
            row.names = F, quote = F, col.names = F)

# prior matrix: pwm scores to bias variable selection toward pairs supported by a TFBS
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)

pwm_imputed <- pwm_occurrence
pwm_imputed[is.na(pwm_imputed)] <- 0.5
priors <- data.frame(pwm_imputed)

write.table(priors, file = "data/inferelator_inputs/priors.tsv", sep = '\t', 
            row.names = T, quote = F)
```

Manually added the tab before colnames to get the exact same format as inferelator examples.


Launching the Infereltaor with stars-lasso without TFA, to get the most comparable outputs:

```{python}

from inferelator import inferelator_workflow, MPControl

n_cores_local = 10
local_engine = True

# Multiprocessing needs to be protected with the if __name__ == 'main' pragma
if __name__ == '__main__' and local_engine:
    MPControl.set_multiprocess_engine("multiprocessing")
    MPControl.client.set_processes(n_cores_local)
    MPControl.connect()
      
worker = inferelator_workflow(regression="elasticnet", workflow="tfa")

worker.set_file_paths(input_dir="data/inferelator_inputs/",
                      output_dir="data/inferelator_outputs",
                      expression_matrix_file="expression.tsv",
                      tf_names_file="tf_names.tsv",
                      priors_file="priors.tsv")
                      
worker.set_run_parameters(num_bootstraps=2, random_seed=42)
worker.set_network_data_flags(use_no_gold_standard=True)
network_result = worker.run()
```

