---
title: "Comparing DIOgene to other existing implementations for integrative GRN inference"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 10)

source('inference_functions/weightedRF.R')
source('inference_functions/weightedLASSO.R')
source("inference_functions/mLASSO_StARS.R")
source('inference_functions/evaluateNetwork.R')
source('inference_functions/data_integration_optimization.R')

library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(ggVennDiagram)
library(ComplexHeatmap)
library(clusterProfiler)
library(circlize)
```

Data loading : 

```{r}
# building expression
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
# x <- t(counts)[,genes]

# building prior (for three values of alpha: 0, 0.5, 1)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)

alphas <- c(0, 0.5, 1)

nCores <- 45
```

# Non linear case: comparison to iRafNet

Differences between iRafNet and weightedRF:

+ Importance metric (Node Purity vs permutation approach)
+ iRafNet has to scale the features to get appropriate feature importance


```{r, eval = F}
# orifinal iRafNet, simply with the MSE having been added as an output
# library(iRafNet)


results <- NULL
for(alpha in alphas){
  # computing models for iRafNet and weightedRF
  importances_iRafNet <- weightedRF_inference(counts, genes, tfs, alpha = alpha, nTrees = 2000,
                                       importance = "IncNodePurity", pwm_occurrence = pwm_occurrence, 
                                       nCores = nCores)

  importances_wRF <- weightedRF_inference(counts, genes, tfs, alpha = alpha, nTrees = 2000,
                                         pwm_occurrence = pwm_occurrence, 
                                         nCores = nCores)
  
  # building sparse GRNs and evaluating them against DAP-Seq
  pr_curve_iRafNet <- evaluate_fully_connected(importances_iRafNet[tfs,], 
                                               pwm_occurrence = pwm_occurrence,
                            input_genes = genes, input_tfs = tfs, 
                                           validation = c("DAPSeq"),
                                           nCores =nCores) %>%
    mutate(alpha = alpha, method = "iRafNet", 
           mse = median(importances_iRafNet["mse",], na.rm = T))
  
  pr_curve_wRF <- evaluate_fully_connected(importances_wRF[tfs,], 
                                           pwm_occurrence = pwm_occurrence,
                            input_genes = genes, input_tfs = tfs, 
                                           validation = c("DAPSeq"),
                                           nCores =nCores) %>%
    mutate(alpha = alpha, method = "weightedRF",
           mse = median(importances_wRF["mse",], na.rm = T))
  
  if(is.null(results)) results <- rbind.data.frame(pr_curve_iRafNet, pr_curve_wRF)
  else results <- rbind(results, rbind.data.frame(pr_curve_iRafNet, pr_curve_wRF))
}

save(results, file = "results/irafnet_weightedRF_comparison.rdata")


```

Showing the results:

```{r}

load("results/irafnet_weightedRF_comparison.rdata")

plot_rf <- (results %>%
  ggplot(aes(y=precision, x = recall, color = method)) +
  geom_point() + geom_line()+
  ggh4x::facet_nested_wrap(vars(paste(alpha)))+
  theme_bw() )/(
results %>%
  group_by(alpha, method) %>%
  distinct(mse, .keep_all = T) %>%
  ggplot(aes(y=mse, x =method, fill = method)) +
  geom_bar(stat = "identity", width = 0.25) + 
  ggh4x::facet_nested_wrap(vars(paste(alpha)))+
  theme_bw())+
results %>%
  filter(network_name < 0.075)%>%
  ggplot(aes(y=pwm_supprt, x =alpha, color = method)) +
  geom_point() + 
  geom_line(aes(group = interaction(network_name, method), 
                alpha = (1-as.numeric(network_name))*0.5+0.5), show.legend = F)+
  # ggh4x::facet_nested_wrap(vars(paste(network_name)), ncol = 5)+
  theme_bw()

# results %>%
#   filter(network_name < 0.1)%>%
#   ggplot(aes(y=pwm_supprt, x =alpha, color = method)) +
#   geom_point() + geom_line()+
#   ggh4x::facet_nested_wrap(vars(paste(network_name)), ncol = 5)+
#   theme_bw()

ggexport(plot_rf, filename = "results/comparisons_to_existing_methods/non_linear_comparison.pdf", 
         width = 8, height = 8)


```


# Linear case : closest implementation to mLASSO-stars

Differences between weightedLASSO and mLASSO-StARS:

+ Glm family is Poisson for weightedLASSO, and Gaussian (following log transform) for mLASSO-StARS 
+ Sparsity selection is done by cv.glmnet in weightedLASSO, and by lasso.stars in mLASSO-StARS
+ The importance metric is MSE increase on OOB for weightedLASSO, and selection frequency + partial correlation for mLASSO-StARS.

```{r}
results <- NULL
for(alpha in alphas){
  # computing models for mLASSO_stars and weightedLASSO
  importances_mLASSO_stars <- mLASSO_stars_inference(counts, genes, tfs, alpha = alpha, 
                                                     N_boot = 5, N_stars_ss = 5,
                                                     pwm_occurrence = pwm_occurrence, 
                                                     nCores = nCores)

  importances_wLASSO <- weightedLASSO_inference(counts, genes, tfs, alpha = alpha, 
                                                N = 25, pwm_occurrence = pwm_occurrence, 
                                                nCores = nCores)
  
  # building sparse GRNs and evaluating them against DAP-Seq
  pr_curve_mLASSO_stars <- evaluate_fully_connected(importances_mLASSO_stars[tfs,], 
                                               pwm_occurrence = pwm_occurrence,
                            input_genes = genes, input_tfs = tfs, 
                                           validation = c("DAPSeq"),
                                           nCores =10) %>%
    mutate(alpha = alpha, method = "mLASSO_stars", 
           mse = median(importances_mLASSO_stars["mse",], na.rm = T))
  
  pr_curve_wLASSO <- evaluate_fully_connected(importances_wLASSO[tfs,], 
                                           pwm_occurrence = pwm_occurrence,
                            input_genes = genes, input_tfs = tfs, 
                                           validation = c("DAPSeq"),
                                           nCores =10) %>%
    mutate(alpha = alpha, method = "weightedLASSO",
           mse = median(importances_wLASSO["mse",], na.rm = T))
  
  if(is.null(results)) results <- rbind.data.frame(pr_curve_wLASSO, pr_curve_mLASSO_stars)
  else results <- rbind(results, rbind.data.frame(pr_curve_wLASSO, pr_curve_mLASSO_stars))
}

save(results, file = "results/mLASSO_stars_weightedLASSO_comparison.rdata")


```

Results:

```{r}
load(file = "results/mLASSO_stars_weightedLASSO_comparison.rdata")


plot_lasso <- (results %>%
  ggplot(aes(y=precision, x = recall, color = method)) +
  geom_point() + geom_line()+
  ggh4x::facet_nested_wrap(vars(paste(alpha)))+
  theme_bw() )/(
results %>%
  group_by(alpha, method) %>%
  distinct(mse, .keep_all = T) %>%
  ggplot(aes(y=mse, x =method, fill = method)) +
  geom_bar(stat = "identity", width = 0.25) + 
  ggh4x::facet_nested_wrap(vars(paste(alpha)))+
  theme_bw())+
results %>%
  filter(network_name < 0.075)%>%
  ggplot(aes(y=pwm_supprt, x =alpha, color = method)) +
  geom_point() + 
  geom_line(aes(group = interaction(network_name, method), 
                alpha = (1-as.numeric(network_name))*0.5+0.5 ), show.legend = F)+
  # ggh4x::facet_nested_wrap(vars(paste(network_name)), ncol = 5)+
  theme_bw()

ggexport(plot_lasso, filename = "results/comparisons_to_existing_methods/linear_comparison.pdf", width = 8, height = 8)
```






<!-- comparison to stars-lasso in the Inferlator? -->

<!-- Generating input tsv files to feed the Inferelator worker: -->

<!-- ```{r} -->
<!-- # expression file :  -->
<!-- load('rdata/inference_input_N_response_varala.rdata') -->
<!-- genes <- input_data$grouped_genes; length(genes) -->
<!-- tfs <- input_data$grouped_regressors; length(tfs) -->
<!-- counts <- input_data$counts; dim(counts) -->

<!-- # expression should be log transformed as the model is  -->
<!-- expression <- log10(counts + 0.75) %>% -->
<!--   as.data.frame()  -->
<!-- options(OutDec=".") -->
<!-- # %>% -->
<!-- #   rownames_to_column("gene") -->

<!-- write.table(expression, file = "data/inferelator_inputs/expression.tsv",  -->
<!--             sep = '\t', row.names = T, quote = F) -->

<!-- # names of the TFs -->
<!-- write.table(tfs, file = "data/inferelator_inputs/tf_names.tsv", sep = '\t',  -->
<!--             row.names = F, quote = F, col.names = F) -->

<!-- # prior matrix: pwm scores to bias variable selection toward pairs supported by a TFBS -->
<!-- load("rdata/pwm_occurrences_N_response_varala.rdata") -->
<!-- dim(pwm_occurrence) -->

<!-- pwm_imputed <- pwm_occurrence -->
<!-- pwm_imputed[is.na(pwm_imputed)] <- 0.5 -->
<!-- priors <- data.frame(pwm_imputed) -->

<!-- write.table(priors, file = "data/inferelator_inputs/priors.tsv", sep = '\t',  -->
<!--             row.names = T, quote = F) -->
<!-- ``` -->

<!-- Manually added the tab before colnames to get the exact same format as inferelator examples. -->


<!-- Launching the Infereltaor with stars-lasso without TFA, to get the most comparable outputs: -->

<!-- ```{python} -->

<!-- from inferelator import inferelator_workflow, MPControl -->

<!-- n_cores_local = 10 -->
<!-- local_engine = True -->

<!-- # Multiprocessing needs to be protected with the if __name__ == 'main' pragma -->
<!-- if __name__ == '__main__' and local_engine: -->
<!--     MPControl.set_multiprocess_engine("multiprocessing") -->
<!--     MPControl.client.set_processes(n_cores_local) -->
<!--     MPControl.connect() -->

<!-- worker = inferelator_workflow(regression="elasticnet", workflow="tfa") -->

<!-- worker.set_file_paths(input_dir="data/inferelator_inputs/", -->
<!--                       output_dir="data/inferelator_outputs", -->
<!--                       expression_matrix_file="expression.tsv", -->
<!--                       tf_names_file="tf_names.tsv", -->
<!--                       priors_file="priors.tsv") -->

<!-- worker.set_run_parameters(num_bootstraps=2, random_seed=42) -->
<!-- worker.set_network_data_flags(use_no_gold_standard=True) -->
<!-- network_result = worker.run() -->
<!-- ``` -->

