---
title: "Comparing DIOgene to other existing implementations for integrative GRN inference"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 10)

source('inference_functions/weightedRF.R')
source('inference_functions/weightedLASSO.R')
source("inference_functions/mLASSO_StARS.R")
source('inference_functions/evaluateNetwork.R')
source('inference_functions/data_integration_optimization.R')

library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(ggVennDiagram)
library(ComplexHeatmap)
library(clusterProfiler)
library(circlize)
library(MESS)
```

Data loading : 

```{r}
# building expression
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
# x <- t(counts)[,genes]

# building prior (for three values of alpha: 0, 0.5, 1)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)

alphas <- c(0, 0.5, 1)

nCores <- 45
N_rep = 3
```

# Non linear case: comparison to iRafNet

Differences between iRafNet and weightedRF:

+ Importance metric is Node Purity in iRafNet, vs MSE increase after permutations on OOB for weightedRF
+ iRafNet has to scale the features to get appropriate feature importance


```{r, eval = F}
# orifinal iRafNet, simply with the MSE having been added as an output
# library(iRafNet)


results <- NULL
for(alpha in alphas){
  for(rep in 1:N_rep){
    # computing models for iRafNet and weightedRF
    importances_iRafNet <- weightedRF_inference(counts, genes, tfs, alpha = alpha, nTrees = 2000,
                                         importance = "IncNodePurity", pwm_occurrence = pwm_occurrence, 
                                         nCores = nCores)
  
    importances_wRF <- weightedRF_inference(counts, genes, tfs, alpha = alpha, nTrees = 2000,
                                           pwm_occurrence = pwm_occurrence, 
                                           nCores = nCores)
    
    # building sparse GRNs and evaluating them against DAP-Seq
    pr_curve_iRafNet <- evaluate_fully_connected(importances_iRafNet[tfs,], 
                                                 pwm_occurrence = pwm_occurrence,
                              input_genes = genes, input_tfs = tfs, 
                                             validation = c("DAPSeq"),
                                             nCores =nCores) %>%
      mutate(alpha = alpha, method = "iRafNet", rep = rep,
             mse = median(importances_iRafNet["mse",], na.rm = T),
             aupr = MESS::auc(recall, precision, type = 'spline'))
    
    pr_curve_wRF <- evaluate_fully_connected(importances_wRF[tfs,], 
                                             pwm_occurrence = pwm_occurrence,
                              input_genes = genes, input_tfs = tfs, 
                                             validation = c("DAPSeq"),
                                             nCores =nCores) %>%
      mutate(alpha = alpha, method = "weightedRF", rep = rep,
             mse = median(importances_wRF["mse",], na.rm = T),
             aupr = MESS::auc(recall, precision, type = 'spline'))
    
    if(is.null(results)) results <- rbind.data.frame(pr_curve_iRafNet, pr_curve_wRF)
    else results <- rbind(results, rbind.data.frame(pr_curve_iRafNet, pr_curve_wRF))
  }
}

save(results, file = "results/irafnet_weightedRF_comparison_reps.rdata")


```

Showing the results:

```{r, fig.width=20}

load("results/irafnet_weightedRF_comparison_reps.rdata")

plot_rf <- (results %>%
              group_by(method, alpha) %>%
              mutate(aupr = mean(aupr, na.rm = T)) %>%
              mutate(aupr_pos = ifelse(method == "weightedRF", 0.75, 0.25)) %>%
  ggplot(aes(y=precision, x = recall, color = method)) +
  geom_point(size = 0.5) + geom_smooth(method = "gam", aes(fill = method), alpha = 0.2)+
  ggh4x::facet_nested_wrap(vars(paste("alpha =", alpha)), nest_line = T)+
    geom_text(aes(y=0.5, x=aupr_pos, label = paste("AUPR =",round(aupr, 2))), show.legend = F) )/(
results %>%
  group_by(alpha, method) %>%
  mutate(mean_mse = mean(mse, na.rm = T)) %>%
  ggplot(aes(y=mean_mse, x =method, fill = method)) +
  geom_segment(aes(xend = method, y = 0, yend = mean_mse, color = method),size = 2.5) + 
  geom_point(aes(y=mse), size = 0.85, color = "black")+
  ggh4x::facet_nested_wrap(vars(paste("alpha =", alpha)), nest_line = T));plot_rf


plot_rf <- plot_rf & scale_color_manual(values = c("#668877", "#70AD47"))& 
  scale_fill_manual(values = c("#668877", "#70AD47"))&
  theme_bw() & theme(strip.background = element_blank())

# results %>%
#   filter(network_name < 0.075)%>%
#   ggplot(aes(y=pwm_supprt, x =alpha, color = method)) +
#   geom_point() + 
#   geom_line(aes(group = interaction(network_name, method), 
#                 alpha = (1-as.numeric(network_name))*0.5+0.5), show.legend = F)+
#   # ggh4x::facet_nested_wrap(vars(paste(network_name)), ncol = 5)+
#   theme_bw()

# results %>%
#   filter(network_name < 0.1)%>%
#   ggplot(aes(y=pwm_supprt, x =alpha, color = method)) +
#   geom_point() + geom_line()+
#   ggh4x::facet_nested_wrap(vars(paste(network_name)), ncol = 5)+
#   theme_bw()

ggexport(plot_rf, filename = "results/comparisons_to_existing_methods/non_linear_comparison.pdf", 
         width = 10, height = 6)


```


# Linear case : closest implementation to mLASSO-stars

Differences between weightedLASSO and mLASSO-StARS:

+ Glm family is Poisson for weightedLASSO, and Gaussian (following log transform) for mLASSO-StARS 
+ Sparsity selection is done by cv.glmnet in weightedLASSO (lambda 1se), and by the lasso.stars R package in mLASSO-StARS
+ The importance metric is the MSE increase after permutations on OOB for weightedLASSO, and selection frequency + partial correlation for mLASSO-StARS.

```{r}
results <- NULL
for(alpha in alphas){
  
  for(rep in 1:N_rep){
    # computing models for mLASSO_stars and weightedLASSO
    importances_mLASSO_stars <- mLASSO_stars_inference(counts, genes, tfs, alpha = alpha, 
                                                       N_boot = 5, N_stars_ss = 5,
                                                       stars.thresh = 0.05,
                                                       pwm_occurrence = pwm_occurrence, 
                                                       nCores = nCores)
  
    importances_wLASSO <- weightedLASSO_inference(counts, genes, tfs, alpha = alpha, 
                                                  N = 25, pwm_occurrence = pwm_occurrence, 
                                                  nCores = nCores)
    
    # building sparse GRNs and evaluating them against DAP-Seq
    pr_curve_mLASSO_stars <- evaluate_fully_connected(importances_mLASSO_stars[tfs,], 
                                                 pwm_occurrence = pwm_occurrence,
                              input_genes = genes, input_tfs = tfs, 
                                             validation = c("DAPSeq"),
                                             nCores =10) %>%
      mutate(alpha = alpha, method = "mLASSO_stars", rep = rep,
             mse = median(importances_mLASSO_stars["mse",], na.rm = T),
             aupr = MESS::auc(recall, precision, type = 'spline'))
    
    pr_curve_wLASSO <- evaluate_fully_connected(importances_wLASSO[tfs,], 
                                             pwm_occurrence = pwm_occurrence,
                              input_genes = genes, input_tfs = tfs, 
                                             validation = c("DAPSeq"),
                                             nCores =10) %>%
      mutate(alpha = alpha, method = "weightedLASSO", rep = rep,
             mse = median(importances_wLASSO["mse",], na.rm = T),
             aupr = MESS::auc(recall, precision, type = 'spline')) 
    
    
    if(is.null(results)) results <- rbind.data.frame(pr_curve_wLASSO, pr_curve_mLASSO_stars)
    else results <- rbind(results, rbind.data.frame(pr_curve_wLASSO, pr_curve_mLASSO_stars))
  }
}

save(results, file = "results/mLASSO_stars_weightedLASSO_comparison_reps.rdata")


```

Results:

```{r}
load(file = "results/mLASSO_stars_weightedLASSO_comparison_reps.rdata")


plot_lasso <- (results %>%
              group_by(method, alpha) %>%
              mutate(aupr = mean(aupr, na.rm = T)) %>%
              mutate(aupr_pos = ifelse(method == "weightedLASSO", 0.75, 0.25)) %>%
  ggplot(aes(y=precision, x = recall, color = method)) +
  geom_point(size = 0.5) + geom_smooth(method = "gam", aes(fill = method), alpha = 0.2)+
  ggh4x::facet_nested_wrap(vars(paste("alpha =", alpha)), nest_line = T)+
    geom_text(aes(y=0.5, x=aupr_pos, label = paste("AUPR =",round(aupr, 2))), show.legend = F) )/(
results %>%
  group_by(alpha, method) %>%
  mutate(mean_mse = mean(mse, na.rm = T)) %>%
  ggplot(aes(y=mean_mse, x =method, fill = method)) +
  geom_segment(aes(xend = method, y = 0, yend = mean_mse, color = method),size = 2.5) + 
  geom_point(aes(y=mse), size = 0.85, color = "black")+
  ggh4x::facet_nested_wrap(vars(paste("alpha =", alpha)), nest_line = T));plot_lasso


plot_lasso <- plot_lasso & scale_color_manual(values = c("#668877", "#70AD47"))& 
  scale_fill_manual(values = c("#668877", "#70AD47"))&
  theme_bw() & theme(strip.background = element_blank());plot_lasso


ggexport(plot_lasso, filename = "results/comparisons_to_existing_methods/linear_comparison.pdf", width = 10, height = 6)
```

Overlaying the results of DIOgene over these results:


```{r, fig.width = 15}
# loading global alpha results
load(file = "results/mLASSO_stars_weightedLASSO_comparison_reps.rdata")
all_results <- results

load(file = "results/irafnet_weightedRF_comparison_reps.rdata")
all_results <- rbind.data.frame(all_results, results)

# loading DIOgene importance matrices
load("results/rdata/gene_specific/gene_specific_grns_true_sd.rdata")
mats <- mats[str_detect(names(mats),"true")]
names(mats)

# computing DIOgene PR curves
mat_diogene_linear <- apply(simplify2array(mats[str_detect(names(mats), "LASSO")]), 1:2, mean) 
mat_diogene_non_linear <- apply(simplify2array(mats[str_detect(names(mats), "RF")]), 1:2, mean) 


linear_eval <- evaluate_fully_connected(mat_diogene_linear[tfs,],
                                        pwm_occurrence = pwm_occurrence,
                              input_genes = genes, input_tfs = tfs, 
                                             validation = c("DAPSeq"),
                                             nCores =10) %>%
  mutate(aupr = MESS::auc(recall, precision, type = 'spline'),
         mse = median(mat_diogene_linear["mse",]), 
         method = "weightedLASSO", alpha = "DIOgene")

non_linear_eval <- evaluate_fully_connected(mat_diogene_non_linear[tfs,],
                                            pwm_occurrence = pwm_occurrence,
                              input_genes = genes, input_tfs = tfs, 
                                             validation = c("DAPSeq"),
                                             nCores =10)%>%
  mutate(aupr = MESS::auc(recall, precision, type = 'spline'),
         mse = median(mat_diogene_non_linear["mse",]), 
         method = "weightedRF", alpha = "DIOgene")

all_diogene_results <- rbind.data.frame(non_linear_eval, linear_eval) %>%
  mutate(rep = 1, pwm_supprt = NA)


all_results <- all_results %>%
  group_by(method, alpha, rep) %>%
  mutate(aupr_splines = MESS::auc(recall, precision, type = 'spline'),
         aupr_linear = MESS::auc(recall, precision, type = 'linear'),
         Bolstad2 = Bolstad2::sintegral(recall, precision)$int)

plot(y=linear_eval$precision, linear_eval$recall)

# combine all mse and aupr data:
all_results <- rbind.data.frame(all_results, all_diogene_results[,colnames(all_results)])


# computing alternative AUPRs:


mse_max <- 0.3
mse_min <- 0.1

# massive plot:
method_comparison <- (all_results %>%
    mutate(color = ifelse(method == "iRafNet" | method == "mLASSO_stars" , "ref",
                          ifelse((method == "weightedLASSO" | method == "weightedRF") & 
                                   alpha == "DIOgene", "diogene", "weighted")),
           regression = ifelse(str_detect(method, "LASSO"), "Linear case", "Non linear case")) %>%
              group_by(method, alpha) %>%
              mutate(aupr = mean(aupr_linear, na.rm = T))  %>%
      group_by(method, alpha, network_name) %>%
      mutate(mean_precision =mean(precision),
             mean_recall = mean(recall)) %>%
              mutate(aupr_pos = ifelse(str_detect(method, "weighted"), 0.75, 0.25)) %>%
  ggplot(aes(y=precision, x = recall, color = interaction(color, regression))) +
  geom_point(size = 0.35, show.legend = F) + 
    geom_line(aes(y=mean_precision, x = mean_recall,
                  group = interaction(color, regression)), show.legend = T)+
  ggh4x::facet_nested_wrap(vars(regression, paste("alpha =", alpha)), nest_line = T, ncol = 4)+
    theme(legend.position = "none") +
    geom_text(aes(y=0.5, x=aupr_pos, label = paste("AUPR =",round(aupr, 2))), show.legend = F))/ 
(all_results %>%
    mutate(color = ifelse(method == "iRafNet" | method == "mLASSO_stars" , "ref",
                          ifelse((method == "weightedLASSO" | method == "weightedRF") & alpha == "DIOgene", 
                                 "diogene", "weighted")),
           regression = ifelse(str_detect(method, "LASSO"), "Linear case", "Non linear case")) %>%
              group_by(alpha, method) %>%
  mutate(mean_mse = mean(mse, na.rm = T),
         mean_mse_to_plot = ifelse(mean_mse > mse_max, mse_max, mean_mse),
         mse_label_y = mean_mse_to_plot,
         label_nudge = ifelse(mean_mse > mse_max, -0.05, 0.05)) %>%
  ggplot(aes(y=mean_mse, x =method, fill = method)) +
   ylab("Median MSE") + xlab("")+ ylim(c(mse_min,mse_max))+
  geom_segment(aes(xend = method, y = mse_min, yend = mean_mse_to_plot, 
                   color = interaction(color, regression)),size = 2.5) + 
  geom_point(aes(y=mse), size = 0.85, color = "black", show.legend = F)+
   geom_text(aes(label = round(mean_mse, 3), y=0.12, color = interaction(color, regression)), 
             nudge_x = 0.275, fill = "white", show.legend = F)+
  ggh4x::facet_nested_wrap(vars(regression, paste("alpha =", alpha)), 
                           nest_line = T, ncol = 4, scales = "free_x"))  & 
  scale_color_manual(name = "Method",
                     values = c("royalblue", "#668877", "#70AD47", "royalblue", "#668877", "#70AD47"), 
                     labels = c("DIOgene (linear)", "mLASSO-StARS", "weightedLASSO", 
                                "DIOgene (non linear)", "iRafNet", "weightedRF"))& 
  scale_fill_manual(name = "Method",
                     values = c("royalblue", "#668877", "#70AD47", "royalblue", "#668877", "#70AD47"), 
                     labels = c("DIOgene (linear)", "mLASSO-StARS", "weightedLASSO", 
                                "DIOgene (non linear)", "iRafNet", "weightedRF"))& 
  theme_bw() & theme(strip.background = element_blank(), 
                     legend.position = "right")& 
  plot_annotation(tag_levels = "a" );method_comparison

ggexport(method_comparison, filename = "results/comparisons_to_existing_methods/comparison.pdf", 
         width = 11, height = 8)


```

