---
title: "DIOgene tutorial"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

source('inference_functions/DIOgene.R')

```

This document demonstrates the use of DIOgene based on a plant biology case study: the regulatory pathways of *Arabidopsis thaliana*'s roots in response to nitrate (N) induction from [Varala et al., 2018](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE97500).

DIOgene uses as inputs the expression profiles of genes and TFBM information (PWM scores). 
Prior TFBM information was built by searching in the promoters of the nitrate-responsive genes (-1000 bp upstream and 200bp downstream of the TSS) the PWM of the N-responsive regulators found in the [JASPAR database](https://jaspar.genereg.net/). 

## Data import


### Expression data

Import of the expression data and the N-responsive genes and regulators :

```{r}
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
```

### TFBS data


```{r}
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)
```

The following code illustrates how to use DIOgene on a subset of the data (the first 5 genes) and light parameters, for a fast demonstration. The code to reproduce the results of the paper can be found at the end of this notebook.

```{r}
genes <- genes[1:5]
```


# Running DIOgene

## Non-linear model (weightedRF)

```{r}
diogene_results <- DIOgene(counts, genes, tfs, pwm_occurrence, 
                           model = "non-linear", nrep = 5, 
                           nCores = 5, 
                           S = 100)

```


The diogene attribute in the results is the importance matrix between TFs and target genes with a gene specific tuning of alpha.
A final sparse GRN can then be pruned to a specific density with the following function:


```{r}

grn_RF <- weightedRF_network(diogene_results$diogene, 
                             density = 0.005, pwm_occurrence, 
                             genes, tfs)
  
head(grn_RF); dim(grn_RF)
mean(grn_RF$pwm)
```


Finally, the following code can be used to plot the optimal alpha and the MSE trends as a function of EDI for a given gene (here taken randomly):

```{r}
get_opt_alpha_per_gene(gene = sample(genes, size = 1), 
                       mats = diogene_results$mats, 
                       lmses = diogene_results$mse)
```


The resulting GRN can then be mined for biological insights...

## Linear model (weightedLASSO)

The use of DIOgene is very similar in the linear case:

```{r}
diogene_results <- DIOgene(counts, genes, tfs, pwm_occurrence, 
                           model = "linear", nrep = 3, 
                           nCores = 5, 
                           S = 5)

grn_lasso <- weightedLASSO_network(importances_lasso, density = 0.005, 
                                       pwm_occurrence, genes, tfs, decreasing = TRUE)
head(grn_lasso); dim(grn_lasso)
mean(grn_lasso$pwm)
```

# Settings to reproduce the results of the article


These commands and parameters should be used to obtain results like those presented in our article:
(`genes` being the 1426 nitrate responsive genes).


```{r}
diogene_results_non_linear <- DIOgene(counts, genes, tfs, pwm_occurrence, 
                           model = "non-linear", nrep = 100)
diogene_results_linear <- DIOgene(counts, genes, tfs, pwm_occurrence, 
                           model = "linear", nrep = 100)
```

