---
title: "Integrative GRN inference : building GRN with alpha-specific genes"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 10)

source('inference_functions/bRF.R')
source('inference_functions/LASSO-D3S.R')
source('inference_functions/evaluateNetwork.R')
source('inference_functions/MSE.R')
# library(ComplexHeatmap)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(ComplexHeatmap)

```

This document demonstrates the use of the bRF and LASSO-D3S functions for integrative GRN inference.

Those functions infer the regulatory pathways of *Arabidopsis thaliana*'s roots in response to nitrate (N) induction from [Varala et al., 2018](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE97500).

They use as inputs the expression profiles of N-responsive genes and TFBS information. 
Prior TFBS information was built by searching in the promoters of the N-responsive genes the PWM of the N-responsive regulators. 

# Data import

Import of the expression data and the N-responsive genes and regulators :

```{r}
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)

# values of mse per gene per alpha per rep for true and shuffled data
load("results/brf_perumtations_mse_all_genes.rdata")
lmses_rf <- lmses
load("results/lasso_perumtations_mse_all_genes.rdata")
lmses_lasso <- lmses
rm(lmses)

# groups of genes depending on their behavior toward data integration
load("results/clusters_mse_bRF_100permutations.rdata")
positive_genes_rf <- names(clusters_rf[clusters_rf==2])

load("results/clusters_mse_lasso_100permutations.rdata")
positive_genes_lasso <- names(clusters_lasso[clusters_lasso==1])
aphas_per_gene_lasso_mda <- 
  setNames(as.numeric(str_replace(clusters_lasso, "2|no diff", "0")), nm = names(clusters_lasso))

```

# Get optimal alphas per gene

This chunk of code returns the value of alpha where there is a maximal difference between true data and shuffled data in the MSE curves.

```{r, eval=F}
get_optimal_alphas_per_genes <- function(model = "rf", strategy = "min_mse",
                                         alpha_for_other_genes = 0){
  if(model == "rf"){
    lmses = lmses_rf
    positive_genes = positive_genes_rf
  }
  
  if(model == "lasso"){
    lmses = lmses_lasso
    positive_genes = positive_genes_lasso
  }
data <- lmses %>%
rownames_to_column('gene') %>%
  reshape2::melt()%>%
  separate(variable,
           into = c("alpha", "rep", "MSEtype"),
           sep = " ") %>%
  group_by(gene, alpha, MSEtype) %>%
  mutate(mean_mse = mean(value, na.rm = T),
         sd_mse = sd(value, na.rm = T)) %>%
  dplyr::select(mean_mse, sd_mse, gene, alpha, MSEtype)%>%
  distinct() 

data_true <- filter(data, MSEtype=="true_data")
data_perm <- filter(data, MSEtype=="shuffled")

data_true$mean_mse_perm <- data_perm$mean_mse
data_true$sd_mse_perm <- data_perm$sd_mse

if(strategy == "min_mse"){
  alphas_opt <- data_true %>%
  filter(gene %in% positive_genes) %>%
   mutate(mean_mse_diff = mean_mse) %>%
  group_by(gene) %>%
  slice_min(order_by = mean_mse_diff) %>%
  select(gene, alpha) %>%
  column_to_rownames("gene")
}

  if(strategy == "max_div"){
    alphas_opt <- data_true %>%
    filter(gene %in% positive_genes) %>%
    mutate(mean_mse_diff = (mean_mse-mean_mse_perm)/(sd_mse_perm+sd_mse)) %>%
    group_by(gene) %>%
    slice_min(order_by = mean_mse_diff) %>%
    select(gene, alpha) %>%
    column_to_rownames("gene")
  }
  # named vector,
  # adding a value of alpha for genes that do not 
  # benefit from data integration
  all_alphas_per_genes <- c(setNames(as.numeric(alphas_opt$alpha), 
                               rownames(alphas_opt)), 
                          setNames(rep(alpha_for_other_genes, 
                                       length(setdiff(genes, positive_genes))), 
                                  nm=setdiff(genes, positive_genes)))
  
  return(all_alphas_per_genes)
}

alphas_per_genes = list()
for(model in c("rf", "lasso")){
  for(strat in c("min_mse", "max_div")){
    for(alphas_other in c(0,1)){
      alphas_per_genes[[paste0(model, "_", strat, "_", alphas_other)]] <- 
      get_optimal_alphas_per_genes(model = model, strategy = strat, 
                                   alpha_for_other_genes =  alphas_other)
    }
  }
}

save(alphas_per_genes, file = "results/alphas_per_genes.rdata")

```

## see the optimal alpha in the curves

```{r}

load("results/alphas_per_genes.rdata")

draw_gene_mean_sd <- function(gene, model="rf", title = NULL){
  if(model == "rf"){
    lmses = lmses_rf
  }
  
  if(model == "lasso"){
    lmses = lmses_lasso
  }
  
  plot <- lmses[gene, ] %>%
    gather() %>%
    separate(key,
             into = c("alpha", "rep", "MSEtype"),
             sep = " ") %>%
    group_by(alpha, MSEtype) %>%
    mutate(mean_mse = mean(value, na.rm = T),
           sd_mse = sd(value, na.rm = T),
           alpha = as.numeric(alpha)) %>%
    ggplot(aes(
      x = as.numeric(alpha),
      y = value,
      color = MSEtype,
      fill = MSEtype
    )) +ggtitle(paste(gene, title))+ylab("MSE/Var(gene)")+
    geom_ribbon(aes(ymin = mean_mse - sd_mse , 
                    ymax = mean_mse + sd_mse  ), 
                alpha = .4)  +theme_pubr(legend = "top")+
    geom_point(alpha = 0.1) + geom_smooth(se=F)+xlab("alpha") 
  if(model == "rf")
    return(plot+
    geom_vline(xintercept = as.numeric(alphas_per_genes$rf_min_mse_0[gene]))+
    geom_vline(xintercept = as.numeric(alphas_per_genes$rf_max_div_0[gene]), color = "darkred"))
  if(model == "lasso")
    return(plot+
    geom_vline(xintercept = as.numeric(alphas_per_genes$lasso_min_mse_0[gene]))+
    geom_vline(xintercept = as.numeric(alphas_per_genes$lasso_max_div_0[gene]), color = "darkred"))
}

for(gene in sample(genes, 10)){
  print(draw_gene_mean_sd(gene, model = "lasso", title = "lasso")+
          draw_gene_mean_sd(gene, model = "rf", title = "rf"))
}

for(gene in "AT1G08090"){
  print(draw_gene_mean_sd(gene, model = "lasso", title = "lasso")+
          draw_gene_mean_sd(gene, model = "rf", title = "rf"))
}

"AT1G08090" %in% positive_genes_rf
```



```{r, eval=F}


nCores = 40
networks <- list()
densities <- c(0.005, 0.01,0.05,0.075)

for(model in c("lasso")){
  for(strat in c("max_div")){
    for(alphas_other in c(0)){
      for(g in c("all", "positive")){
        for(shuffle in c( F, T)){
          for(rep in 1:10){
            if(g == "all") gene_list <- genes
            else{
              if(model == "rf") gene_list <- positive_genes_rf
              else gene_list <- positive_genes_lasso
            }
            
            print(paste0(model, "-", strat, "-", alphas_other, "-", g, "-", 
                               ifelse(shuffle, "shuffled", "trueData"), "-", rep))
            
            if(model == "rf"){
              mat <- bRF_inference(counts=counts, genes= gene_list, tfs=tfs, 
                          alpha = alphas_per_genes[[paste0(model, "_", strat, "_", alphas_other)]], 
              tf_expression_permutation = shuffle, pwm_occurrence = pwm_occurrence,
              nTrees = 2000, importance = "%IncMSE", nCores = nCores)
              
              for(density in densities){
                networks[[paste0(model, "-", strat, "-", alphas_other, "-", g, "-", 
                               ifelse(shuffle, "shuffled", "trueData"), "-", rep, "-", density)]] <-
                bRF_network(mat, density, pwm_occurrence, gene_list, tfs)
              }
            }
            
            if(model == "lasso"){
              mat <- LASSO.D3S_inference_stableMDA(counts, gene_list, tfs,mda_type = "shuffle",
                                   alpha = aphas_per_gene_lasso_mda, 
                                   N = 100, tf_expression_permutation = shuffle,
                                   int_pwm_noise = 0, robustness = 0.2,
                                   pwm_occurrence = pwm_occurrence,
                                   nCores = nCores)
              
              for(density in densities){
                networks[[paste0(model, "-", strat, "-", alphas_other, "-", g, "-", 
                               ifelse(shuffle, "shuffled", "trueData"), "-", rep, "-", density)]] <-
                LASSO.D3S_network(mat, density, pwm_occurrence, gene_list, tfs, decreasing = T)
              }
            }
          }
        }
      }
    }
  }
}
save(networks, file = "results/gene_specific_alphas_grns_max_div_0others_lasso_shuffle.rdata")

```

```{r, fig.width=10, eval=FALSE}
load("results/gene_specific_alphas_grns_max_div_0others.rdata")
nCores = 20


pos_nets <- networks[str_detect(names(networks), 'positive')]
all_nets <- networks[!str_detect(names(networks), 'positive')]

names(pos_nets)

pos_nets_lasso <- pos_nets[str_detect(names(pos_nets), 'lasso')]
pos_nets_rf <- pos_nets[str_detect(names(pos_nets), 'rf')]

all_nets_lasso <- all_nets[str_detect(names(all_nets), 'lasso')]
all_nets_rf <- all_nets[str_detect(names(all_nets), 'rf')]


val <- rbind.data.frame(evaluate_networks(pos_nets_lasso, nCores = nCores,
                  input_genes = positive_genes_lasso ),
                 evaluate_networks(pos_nets_rf, nCores = nCores,
                  input_genes = positive_genes_rf),
                 evaluate_networks(all_nets_lasso, nCores = nCores,
                  input_genes = genes),
                 evaluate_networks(all_nets_rf, nCores = nCores,
                  input_genes = genes ))

val <- rbind.data.frame(evaluate_networks(pos_nets_lasso, nCores = nCores,
                  input_genes = positive_genes_lasso ),
                 evaluate_networks(all_nets_lasso, nCores = nCores,
                  input_genes = genes))

settings <- c("model", "alphaOpt", "alphaOtherGenes", "genesInNetwork", "dataset", "rep", "density")

val[, settings] <-
  str_split_fixed(val$network_name, '-', length(settings))


val %>%
ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  ggh4x::facet_nested_wrap(vars(model, density, genesInNetwork), 
                           ncol = 8, nest_line = T) + 
  geom_point() + stat_compare_means(label = "p.signif")

val %>%
ggplot(aes(x=dataset, y = recall, color = dataset)) + 
  ggh4x::facet_nested_wrap(vars(model, density, genesInNetwork), scales = "free",
                           ncol = 8, nest_line = T) + 
  geom_point() + stat_compare_means(label = "p.signif")


save(val, file = "results/gene_specific_validation_lasso_mda_shuffle.rdata")



```


## On RF model

```{r}
hist(alphas_per_genes$rf_max_div_0[positive_genes_rf])
```


```{r, fig.width=7, fig.height=4, eval=FALSE}

load("results/gene_specific_validation.rdata")

load("results/100_permutations_bRF_validation_cluster_pos.rdata")
val_rf_pos <- val_conn_pos
load("results/100_permutations_bRF_validation.rdata")
val_rf_all <- val_conn


plot_positive <- val_rf_pos[(val_rf_pos$alpha == "0" | val_rf_pos$alpha == "1") & 
             as.numeric(val_rf_pos$rep) <=10 & val_rf_pos$density == 0.005 &
             val_rf_pos$method == "bRF",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("Global alpha")+ 
  stat_compare_means(label = "p.signif")+
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  geom_hline(yintercept = 0.3807947)+
  val %>%
  filter(density == 0.005 & model == "rf" & genesInNetwork == "positive")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") + 
  geom_hline(yintercept = 0.3807947)+
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific alpha") +
  plot_annotation(title = "Precision of 0.005 density networks for bRF ", 
                  subtitle = "GRNs of positive genes only") +
  plot_layout(widths = c(2,1))


plot_all <- val_rf_all[(val_rf_all$alpha == "0") & 
             as.numeric(val_rf_all$rep) <=10 & val_rf_all$density == 0.005 &
             val_rf_all$method == "bRF",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("No integration")+ 
  stat_compare_means(label = "p.signif")+
  geom_hline(yintercept = 0.331042) + geom_jitter()+
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  val %>%
  filter(density == 0.005 & model == "rf" & genesInNetwork == "all")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") + geom_jitter()+
  geom_hline(yintercept = 0.331042) + 
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific data integration") +
  plot_annotation(title = "Precision against DAP-Seq of 0.005 density networks for bRF ", 
                  subtitle = "GRNs of all genes") +
  plot_layout(widths = c(1,1))


load("results/brf_perumtations_mse_all_genes.rdata")
names(lmses)
lmses %>%
  select(contains())
rownames_to_column('gene') %>%
  reshape2::melt()%>%
  separate(variable,
           into = c("alpha", "rep", "MSEtype"),
           sep = " ") %>%
  group_by(gene, alpha, MSEtype)
  
plot_all

```

### DAP only

```{r, fig.width=7, fig.height=4}
# load("results/100_permutations_bRF_validation.rdata")
# 
# load("results/100_permutations_lasso_validation.rdata")
# load("results/100_permutations_lasso_validation_cluster_pos.rdata")
load("results/gene_specific_validation_dap.rdata")

load("results/100_permutations_bRF_validation_dap_cluster_pos.rdata")
val_rf_pos <- val_dap
load("results/100_permutations_bRF_validation_dap.rdata")
val_rf_all <- val_dap


plot_positive <- val_rf_pos[(val_rf_pos$alpha == "0" | val_rf_pos$alpha == "1") & 
             as.numeric(val_rf_pos$rep) <=10 & val_rf_pos$density == 0.005 &
             val_rf_pos$method == "bRF",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("Global alpha")+ 
  stat_compare_means(label = "p.signif")+
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  geom_hline(yintercept = 0.3547275)+
  val %>%
  filter(density == 0.005 & model == "rf" & genesInNetwork == "positive")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") +
  geom_hline(yintercept = 0.3547275)+
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific alpha") +
  plot_annotation(title = "Precision of 0.005 density networks for bRF ", 
                  subtitle = "GRNs of positive genes only") +
  plot_layout(widths = c(2,1))


plot_all <- val_rf_all[(val_rf_all$alpha == "0") & 
             as.numeric(val_rf_all$rep) <=10 & val_rf_all$density == 0.005 &
             val_rf_all$method == "bRF",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("Global alpha")+ 
  stat_compare_means(label = "p.signif")+
  geom_hline(yintercept = 0.331042) + 
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  val %>%
  filter(density == 0.005 & model == "rf" & genesInNetwork == "all")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") +
  geom_hline(yintercept = 0.331042) + 
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific alpha") +
  plot_annotation(title = "Precision of 0.005 density networks for bRF ", 
                  subtitle = "GRNs of all genes") +
  plot_layout(widths = c(1,1))


plot_all
plot_positive

```



## On LASSO model

```{r}
hist(alphas_per_genes$lasso_max_div_0[positive_genes_lasso])
```


```{r, fig.width=7, fig.height=4, eval=FALSE}
load("results/gene_specific_validation_lasso_mda_shuffle.rdata")

load("results/100_permutations_lasso_validation_cluster_pos_mda_shuffle.rdata")
val_lasso_pos <- val_conn_pos
load("results/100_permutations_lasso_validation_mda_shuffle.rdata")
val_lasso_all <- val_conn


plot_positive <- val_lasso_pos[(val_lasso_pos$alpha == "0" | val_lasso_pos$alpha == "1") & 
             as.numeric(val_lasso_pos$rep) <=10 & val_lasso_pos$density == 0.005 &
             val_lasso_pos$method == "LASSO",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("Global alpha")+ 
  stat_compare_means(label = "p.signif")+
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  geom_hline(yintercept = 0.2921348)+
  val %>%
  filter(density == 0.005 & model == "lasso" & genesInNetwork == "positive")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") +
  geom_hline(yintercept = 0.2921348)+
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific alpha") +
  plot_annotation(title = "Precision of 0.005 density networks for LASSO ", 
                  subtitle = "GRNs of positive genes only") +
  plot_layout(widths = c(2,1))


plot_all <- val_lasso_all[(val_lasso_all$alpha == "0") & 
             as.numeric(val_lasso_all$rep) <=10 & val_lasso_all$density == 0.005 &
             val_lasso_all$method == "LASSO",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("Global alpha")+ 
  stat_compare_means(label = "p.signif")+
  geom_hline(yintercept = 0.3451811) + 
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  val %>%
  filter(density == 0.005 & model == "lasso" & genesInNetwork == "all")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") +
  geom_hline(yintercept = 0.3451811) + 
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific alpha") +
  plot_annotation(title = "Precision of 0.005 density networks for LASSO ", 
                  subtitle = "GRNs of all genes") +
  plot_layout(widths = c(1,1))


plot_all
plot_positive

```



### DAP only


```{r, fig.width=7, fig.height=4}
# load("results/100_permutations_bRF_validation.rdata")
# 
# load("results/100_permutations_lasso_validation.rdata")
# load("results/100_permutations_lasso_validation_cluster_pos.rdata")
load("results/gene_specific_validation_lasso_mda_shuffle.rdata")

load("results/100_permutations_lasso_validation_dap_cluster_pos_mda_shuffle.rdata")
val_lasso_pos <- val_dap

load("results/100_permutations_lasso_validation_dap_mda_shuffle.rdata")
val_lasso_all <- val_dap

plot_positive <- val_lasso_pos[(val_rf_pos$alpha == "0" | val_lasso_pos$alpha == "1") & 
             as.numeric(val_lasso_pos$rep) <=10 & val_lasso_pos$density == 0.005 &
             val_lasso_pos$method == "LASSO",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("Global alpha")+ 
  stat_compare_means(label = "p.signif")+
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  geom_hline(yintercept = 0.3337577)+
  val %>%
  filter(density == 0.005 & model == "lasso" & genesInNetwork == "positive")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") +
  geom_hline(yintercept = 0.3337577)+
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific alpha") +
  plot_annotation(title = "Precision of 0.005 density networks for lasso ", 
                  subtitle = "GRNs of positive genes only") +
  plot_layout(widths = c(2,1))


plot_all <- val_lasso_all[(val_lasso_all$alpha == "0") & 
             as.numeric(val_lasso_all$rep) <=10 & val_rf_all$density == 0.005 &
             val_lasso_all$method == "LASSO",] %>%
  ggplot(aes(x=dataset, color = dataset, y = precision)) + 
  ggh4x::facet_nested_wrap(vars(alpha), 
                           ncol = 8, nest_line = T) +
  geom_boxplot(outlier.alpha = 0)+ ggtitle("Global alpha")+ 
  stat_compare_means(label = "p.signif")+
  geom_hline(yintercept = 0.331042) + 
  ylim(c(0.2, 0.5)) + theme_pubr(legend = "none")+
  theme(strip.background = element_blank())+
  val %>%
  filter(density == 0.005 & model == "lasso" & genesInNetwork == "all")%>%
  ggplot(aes(x=dataset, y = precision, color = dataset)) + 
  stat_compare_means(label = "p.signif") + 
  ylim(c(0.2, 0.5)) + stat_compare_means(label = "p.signif")+
  theme_pubr(legend = "none") +
  geom_hline(yintercept = 0.331042) + 
  geom_boxplot(outlier.alpha = 0) + ggtitle("Gene-specific alpha") +
  plot_annotation(title = "Precision of 0.005 density networks for lasso ", 
                  subtitle = "GRNs of all genes") +
  plot_layout(widths = c(1,1))


plot_all
plot_positive

```


