---
title: "Integrative GRN inference via LASSO-D3S: permuting the expression and motifs of TFs: effect on precision and recall"
output: 
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, fig.width = 11, fig.height = 10)

source('inference_functions/LASSO-D3S.R')
source('inference_functions/evaluateNetwork.R')
source('inference_functions/MSE.R')
# library(ComplexHeatmap)
library(ggplot2)
library(tidyverse)
library(ggpubr)
library(patchwork)
# library(ComplexHeatmap)

theme_set(theme_pubr())
```

This document demonstrates the use of the bRF and LASSO-D3S functions for integrative GRN inference.

Those functions infer the regulatory pathways of *Arabidopsis thaliana*'s roots in response to nitrate (N) induction from [Varala et al., 2018](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE97500).

They use as inputs the expression profiles of N-responsive genes and TFBS information. 
Prior TFBS information was built by searching in the promoters of the N-responsive genes the PWM of the N-responsive regulators. 

# Data import

Import of the expression data and the N-responsive genes and regulators :

```{r}
load('rdata/inference_input_N_response_varala.rdata')
genes <- input_data$grouped_genes; length(genes)
tfs <- input_data$grouped_regressors; length(tfs)
counts <- input_data$counts; dim(counts)
load("rdata/pwm_occurrences_N_response_varala.rdata")
dim(pwm_occurrence)


ALPHAS <- c(seq(0,0.8, by = 0.2), 0.9,1)

ALPHAS <- c(0.1,0.3,0.5, 0.7)

# subset <- sample(genes, replace = F, size = 20)
subset <- genes
```


## Lauching the permutations and the true data inferences 100 times 

```{r, eval=FALSE}
nCores = 45
mats <- list()

nrep <- 100
for(alpha in ALPHAS){ # exploring PWM integration strength
  for(rep in 1:nrep){ # exploring inherent variability
    
    
    
    mat_lasso <- LASSO.D3S_inference_stableMDA(counts, genes, tfs,
                                     alpha = alpha, N = 100, robustness = 0.2,
                                     int_pwm_noise = 0, mda_type = "zero",
                                     pwm_occurrence = pwm_occurrence,
                                     nCores = nCores)
    
    mat_lasso_perm <- LASSO.D3S_inference_stableMDA(counts, genes, tfs,
                                     alpha = alpha, N = 100, robustness = 0.2,
                                     tf_expression_permutation = TRUE,
                                     int_pwm_noise = 0, mda_type = "zero",
                                     pwm_occurrence = pwm_occurrence,
                                     nCores = nCores)
    
    mats[[paste0("LASSO_", as.character(alpha),  '_trueData_', rep)]] <- mat_lasso
    mats[[paste0("LASSO_", as.character(alpha),  '_shuffled_', rep)]] <- mat_lasso_perm
    
  }
}



save(mats, file = "results/100_permutations_lasso_importances_mda_shuffle.rdata")
```


```{r}

# 
mean(edges$LASSO_0.9_trueData_1_0.005$pwm)
# 
# hist(mat_lasso, breaks=100)
# 
# 
grn0 <- LASSO.D3S_network(mat_lasso, density = 0.005, pwm_occurrence, genes, tfs, decreasing = T)
mean(grn0$pwm)
# 
evaluate_network(grn0, validation = c("DAPSeq"), input_genes = genes, input_tfs = tfs)[c("tpr", "recall")]
# 
# tf <- "AT3G11280"
# target <- "AT4G32480"

edges <- lapply(edges, na.omit)

```

# Thresholding the networks

```{r, eval=FALSE}
# load( "results/100_permutations_lasso_importances_freq.rdata")
edges <- list()
densities <- c(0.005, 0.01,0.05,0.075)

for(name in names(mats)){ # exploring importance threshold stringency
      for(density in densities){
        edges[[paste0(name, '_', density)]] <-
        LASSO.D3S_network(mats[[name]], density = density, pwm_occurrence, genes, tfs, decreasing = TRUE)
      }
}

save(edges, file = "results/100_permutations_lasso_edges_mda_shuffle.rdata")

```


# Thresholding the networks for positive genes

```{r, eval=FALSE}
load("results/clusters_mse_lasso_100permutations.rdata")
positive_genes <- names(clusters_lasso[clusters_lasso==1])

edges <- list()
densities <- c(0.005, 0.01,0.05,0.075)

for(name in names(mats)){ # exploring importance threshold stringency
      for(density in densities){
        edges[[paste0(name, '_', density)]] <-
        LASSO.D3S_network(mats[[name]][,positive_genes], density = density, pwm_occurrence, positive_genes, tfs, decreasing = TRUE)
      }
}

save(edges, file = "results/100_permutations_lasso_edges_positive_same_density_mda_shuffle.rdata")
```



```{r, eval=FALSE}
load( "results/100_permutations_lasso_importances_freq.rdata")

edges <- list()
densities <- c(0.005, 0.01,0.05,0.075)

pwm_imputed <- pwm_occurrence
pwm_imputed[is.na(pwm_imputed)] <- 0.5
  

## alternative that works for alpha = 1
for(name in names(mats)){ # exploring importance threshold stringency
      for(density in densities){
        l <- setNames(lapply(genes, function(gene){names(which(rank(-mats[[name]][,gene]) < density*length(tfs)))}), genes)
        edges[[paste0(name, '_', density)]] <-
        plyr::ldply(l, rbind)%>%
          rename(to = `.id`) %>%
          pivot_longer( -c("to")) %>%
          select(-name) %>%
          rename(from = value)%>%
          mutate(pwm = pwm_imputed[cbind(to, from)])
      }
}


save(edges, file = "results/100_permutations_lasso_edges_gene_density_mda0.rdata")
names(edges)

edges <- edges[!str_detect(names(edges), "LASSO_1_")]
```


```{r}

color_palette = c("#C12131", "#EC5D2F", "#FE945C", "#FFC08E" )
prettyZero <- function(l){
    max.decimals = max(nchar(str_extract(l, "\\.[0-9]+")), na.rm = T)-1
    lnew = formatC(l, replace.zero = T, zero.print = "0",
        digits = max.decimals, format = "f", preserve.width=T)
    return(lnew)
}

settings <- c("method", "alpha", "dataset","rep", "density")
```

## PWM support

```{r, fig.width=10}
# load("results/100_permutations_lasso_edges_freq.rdata")
# 



# number of edges per network
nrows <-
  data.frame(alpha_rep = names(unlist(lapply(edges, FUN = nrow))),
             n_edges = unlist(lapply(edges, FUN = nrow)))

nrows[, settings] <-
  str_split_fixed(nrows$alpha_rep, '_', length(settings))

edges_num <- lapply(edges, function(df)
  df[sapply(df, is.numeric)])

pwm_support <-
  data.frame(alpha_rep = names(unlist(lapply(edges_num, FUN = nrow))),
             pwm = unlist(lapply(edges_num, FUN = colMeans, na.rm=TRUE)))
pwm_support[, settings] <-
  str_split_fixed(pwm_support$alpha_rep, '_', length(settings))

pwm_support %>%
  left_join(nrows, by = settings) %>%
  mutate(alpha = as.numeric(alpha),
         density_label = paste(density, ':', n_edges, 'edges')) %>%
  ggplot(aes(color = density_label, x = alpha, y = pwm)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) +
  geom_point() +
  geom_smooth(aes(fill = density_label, linetype =dataset) , alpha = 0.1) +
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.position = "top",
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12)
  ) +
  xlab(expression(alpha)) + ylab("Mean PWM score in selected edges") +
  ggtitle("Average PWM support of inferred edges") +
  guides(color = guide_legend(nrow = 10, byrow = TRUE),
         fill = guide_legend(nrow = 10, byrow = TRUE)) +
  ylab(expression(paste("mean(", pi[tr], ")"))) +
  scale_x_continuous(labels = prettyZero)

```

## Computing validation metrics 

### On all-genes networks

```{r, eval=FALSE}
val_conn <-
  evaluate_networks(
    edges,
    validation = c("TARGET", "CHIPSeq", "DAPSeq"),
    nCores = 25,
    input_genes = genes,
    input_tfs = tfs
  )

val_conn[, settings] <-
  str_split_fixed(val_conn$network_name, '_', length(settings))

save(val_conn, file = "results/100_permutations_lasso_validation_mda_shuffle.rdata")

#
val_chip <-
  evaluate_networks(
    edges,
    validation = c("CHIPSeq"),
    nCores = 25,
    input_genes = genes,
    input_tfs = tfs
  )

val_chip[, settings] <-
  str_split_fixed(val_chip$network_name, '_', length(settings))
save(val_chip, file = "results/100_permutations_lasso_validation_chip_freq.rdata")

val_target <-
  evaluate_networks(
    edges,
    validation = c("TARGET"),
    nCores = 25,
    input_genes = genes,
    input_tfs = tfs
  )

val_target[, settings] <-
  str_split_fixed(val_target$network_name, '_', length(settings))
save(val_target, file = "results/100_permutations_lasso_validation_target_freq.rdata")


val_dap <-
  evaluate_networks(
    edges,
    validation = c("DAPSeq"),
    nCores = 35,
    input_genes = genes,
    input_tfs = tfs
  )

val_dap[, settings] <-
  str_split_fixed(val_dap$network_name, '_', length(settings))

save(val_dap, file = "results/100_permutations_lasso_validation_dap_mda_shuffle.rdata")


```


### topology of the all-genes networks

```{r, fig.width=20, fig.height=7}
get_nTF <- function(net){
  return(mean(table(net$to)))
}

get_TF <- function(net){
  return(length(unique(net$from)))
}

get_nGenes <- function(net){
  return(length(unique(net$to)))
}

# subset <- edges[as.numeric(str_split_fixed(names(edges), '_', 5)[,4])<6]

topo <- data.frame(alpha_rep = names(edges),
                n_TFs = unlist(lapply(edges, FUN = get_nTF)),
                TFs = unlist(lapply(edges, FUN = get_TF)),
                n_genes = unlist(lapply(edges, FUN = get_nGenes)))

topo[, settings] <- str_split_fixed(topo$alpha_rep,'_', length(settings))

data_topo <- topo %>%  
  left_join(nrows, by = settings) %>%
  mutate(alpha = as.numeric(alpha)) %>%
  mutate(density_label = paste(density, ':', n_edges, 'edges')) %>%
  group_by(alpha, density, dataset) %>%
  mutate(mean_n_TFs = mean(n_TFs, na.rm=T),
            sd_n_TFs = sd(n_TFs, na.rm=T),
            mean_TFs = mean(TFs, na.rm=T),
            sd_TFs = sd(TFs, na.rm=T),
            mean_nGenes = mean(n_genes, na.rm=T),
            sd_nGenes = sd(n_genes, na.rm=T)) 

regs_per_genes <- data_topo %>%
  ggplot(aes(color = dataset, fill = dataset,x = alpha, y = n_TFs)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) + geom_point(alpha = 0.2) +
  geom_smooth(se = F) +
  geom_ribbon(aes(ymin = mean_n_TFs - sd_n_TFs , 
                    ymax = mean_n_TFs + sd_n_TFs  ), 
                alpha = .4)+
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'none'
  ) +
  scale_x_continuous(labels = prettyZero) +
  xlab(expression(alpha)) + ylab("Average number of regulator per gene") + 
                 ggtitle("Average number of regulator per gene")

regs_in_network <- data_topo %>%
  ggplot(aes(color = dataset, fill = dataset, x = alpha, y = TFs)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) + geom_point(alpha = 0.2) +
  geom_ribbon(aes(ymin = mean_TFs - sd_TFs , 
                    ymax = mean_TFs + sd_TFs  ), 
                alpha = .4)+
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'top'
  ) +
  scale_x_continuous(labels = prettyZero) +
  xlab(expression(alpha))+ ylab("Number of distinct regulators") + 
                 ggtitle("Number of distinct regulators") 


targets_in_network <- data_topo %>%
  ggplot(aes(color = dataset, fill = dataset,x = alpha, y = n_genes)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) + 
  geom_point(alpha = 0.2) +
  geom_smooth(se = F) +
  geom_ribbon(aes(ymin = mean_nGenes - sd_nGenes , 
                    ymax = mean_nGenes + sd_nGenes  ), 
                alpha = .4)+
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'top'
  ) +
  scale_x_continuous(labels = prettyZero) +
  xlab(expression(alpha))+ ylab("Number of distinct targets") + 
                 ggtitle("Number of distinct targets") 


figure <- targets_in_network+regs_in_network+regs_per_genes + plot_layout(guides = "collect") + plot_annotation(tag_levels = 'a');figure
```

### On positive networks

```{r, fig.width=10}
load("results/100_permutations_lasso_edges_positive_same_density_mda_shuffle.rdata")
# number of edges per network
nrows <-
  data.frame(alpha_rep = names(unlist(lapply(edges, FUN = nrow))),
             n_edges = unlist(lapply(edges, FUN = nrow)))

nrows[, settings] <-
  str_split_fixed(nrows$alpha_rep, '_', length(settings))

edges_num <- lapply(edges, function(df)
  df[sapply(df, is.numeric)])

pwm_support <-
  data.frame(alpha_rep = names(unlist(lapply(edges_num, FUN = nrow))),
             pwm = unlist(lapply(edges_num, FUN = colMeans)))
pwm_support[, settings] <-
  str_split_fixed(pwm_support$alpha_rep, '_', length(settings))

pwm_support %>%
  left_join(nrows, by = settings) %>%
  mutate(alpha = as.numeric(alpha),
         density_label = paste(density, ':', n_edges, 'edges')) %>%
  ggplot(aes(color = density_label, x = alpha, y = pwm)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) +
  geom_point() +
  geom_smooth(aes(fill = density_label, linetype =dataset) , alpha = 0.1) +
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.position = "top",
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12)
  ) +
  xlab(expression(alpha)) + ylab("Mean PWM score in selected edges") +
  ggtitle("Average PWM support of inferred edges") +
  guides(color = guide_legend(nrow = 2, byrow = TRUE),
         fill = guide_legend(nrow = 2, byrow = TRUE)) +
  ylab(expression(paste("mean(", pi[tr], ")"))) +
  scale_x_continuous(labels = prettyZero)

```

### topology of the positive genes networks

```{r, fig.width=20, fig.height=7}
get_nTF <- function(net){
  return(mean(table(net$to)))
}

get_TF <- function(net){
  return(length(unique(net$from)))
}

get_nGenes <- function(net){
  return(length(unique(net$to)))
}

# subset <- edges[as.numeric(str_split_fixed(names(edges), '_', 5)[,4])<6]

topo <- data.frame(alpha_rep = names(edges),
                n_TFs = unlist(lapply(edges, FUN = get_nTF)),
                TFs = unlist(lapply(edges, FUN = get_TF)),
                n_genes = unlist(lapply(edges, FUN = get_nGenes)))

topo[, settings] <- str_split_fixed(topo$alpha_rep,'_', length(settings))

data_topo <- topo %>%  
  left_join(nrows, by = settings) %>%
  mutate(alpha = as.numeric(alpha)) %>%
  mutate(density_label = paste(density, ':', n_edges, 'edges')) %>%
  group_by(alpha, density, dataset) %>%
  mutate(mean_n_TFs = mean(n_TFs, na.rm=T),
            sd_n_TFs = sd(n_TFs, na.rm=T),
            mean_TFs = mean(TFs, na.rm=T),
            sd_TFs = sd(TFs, na.rm=T),
            mean_nGenes = mean(n_genes, na.rm=T),
            sd_nGenes = sd(n_genes, na.rm=T)) 

regs_per_genes <- data_topo %>%
  ggplot(aes(color = dataset, fill = dataset,x = alpha, y = n_TFs)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) + geom_point(alpha = 0.2) +
  geom_smooth(se = F) +
  geom_ribbon(aes(ymin = mean_n_TFs - sd_n_TFs , 
                    ymax = mean_n_TFs + sd_n_TFs  ), 
                alpha = .4)+
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'none'
  ) +
  scale_x_continuous(labels = prettyZero) +
  xlab(expression(alpha)) + ylab("Average number of regulator per gene") + 
                 ggtitle("Average number of regulator per gene")

regs_in_network <- data_topo %>%
  ggplot(aes(color = dataset, fill = dataset, x = alpha, y = TFs)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) + geom_point(alpha = 0.2) +
  geom_ribbon(aes(ymin = mean_TFs - sd_TFs , 
                    ymax = mean_TFs + sd_TFs  ), 
                alpha = .4)+
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'top'
  ) +
  scale_x_continuous(labels = prettyZero) +
  xlab(expression(alpha))+ ylab("Number of distinct regulators") + 
                 ggtitle("Number of distinct regulators") 


targets_in_network <- data_topo %>%
  ggplot(aes(color = dataset, fill = dataset,x = alpha, y = n_genes)) +
  ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) + 
  geom_point(alpha = 0.2) +
  geom_smooth(se = F) +
  geom_ribbon(aes(ymin = mean_nGenes - sd_nGenes , 
                    ymax = mean_nGenes + sd_nGenes  ), 
                alpha = .4)+
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'top'
  ) +
  scale_x_continuous(labels = prettyZero) +
  xlab(expression(alpha))+ ylab("Number of distinct targets") + 
                 ggtitle("Number of distinct targets") 


figure <- targets_in_network+regs_in_network+regs_per_genes + plot_layout(guides = "collect") + plot_annotation(tag_levels = 'a');figure
```

Not all densities should be explored with this restriction of genes, as there are too few genes to grant sufficient interactions supported by a prior of 1. We keep only the two smallest densities.

```{r, eval=FALSE}
val_conn_pos <-
  evaluate_networks(
    edges,
    validation = c("TARGET", "CHIPSeq", "DAPSeq"),
    nCores = 25,
    input_genes = genes,
    input_tfs = tfs
  )


val_conn_pos[, settings] <-
  str_split_fixed(val_conn_pos$network_name, '_', length(settings))

save(val_conn_pos, file = "results/100_permutations_lasso_validation_cluster_pos_mda_shuffle.rdata")


# 
# val_chip <-
#   evaluate_networks(
#     edges,
#     validation = c("CHIPSeq"),
#     nCores = 25,
#     input_genes = genes,
#     input_tfs = tfs
#   )
# 
# val_chip[, settings] <-
#   str_split_fixed(val_chip$network_name, '_', length(settings))
# save(val_chip, file = "results/100_permutations_lasso_validation_chip_cluster_pos_freq.rdata")
# 
# 
# val_target <-
#   evaluate_networks(
#     edges,
#     validation = c("TARGET"),
#     nCores = 25,
#     input_genes = genes,
#     input_tfs = tfs
#   )

# val_target[, settings] <-
#   str_split_fixed(val_target$network_name, '_', length(settings))
# save(val_target, file = "results/100_permutations_lasso_validation_target_cluster_pos.rdata")


val_dap <-
  evaluate_networks(
    edges,
    validation = c("DAPSeq"),
    nCores = 25,
    input_genes = genes,
    input_tfs = tfs
  )

val_dap[, settings] <-
  str_split_fixed(val_dap$network_name, '_', length(settings))
save(val_dap, file = "results/100_permutations_lasso_validation_dap_cluster_pos_mda_shuffle.rdata")

```

# Plotting validation


```{r}

draw_precision_recall <- function(validation, data_type){
  data_val <- validation %>%
    group_by(alpha, dataset, density) %>%
    mutate(mean_precision = mean(precision, na.rm = T),
           sd_precision = sd(precision, na.rm = T),
           mean_recall = mean(recall, na.rm = T),
           sd_recall = sd(recall, na.rm = T)) %>%
  dplyr::select(-network_name) %>%
  left_join(nrows, by = settings) %>%
  mutate(alpha = as.numeric(alpha)) %>%
  mutate(density_label = paste(density, ':', n_edges, 'edges')) 

precision <- data_val %>%
    ggplot(aes(
      x = as.numeric(alpha),
      y = precision,
      color = dataset,
      fill = dataset
    ))+
    ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, nest_line = T) + 
    geom_ribbon(aes(ymin = mean_precision - sd_precision , 
                    ymax = mean_precision + sd_precision  ), 
                alpha = .4)  +theme_pubr(legend = "top")+
    geom_point(alpha = 0.1) + geom_smooth(se=F)+xlab("alpha") +
  xlab(expression(alpha)) + ylab("Precision") +
  ggtitle(paste("Precision against", data_type)) +
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'top'
  )+geom_hline(yintercept = 0.3451811)

recall <- data_val%>%
    ggplot(aes(
      x = as.numeric(alpha),
      y = recall,
      color = dataset,
      fill = dataset
    ))+
    ggh4x::facet_nested_wrap(vars(method, density), ncol = 8, 
                             nest_line = T, scales = "free") + 
    geom_ribbon(aes(ymin = mean_recall - sd_recall , 
                    ymax = mean_recall + sd_recall  ), 
                alpha = .4)  +theme_pubr(legend = "none")+
    geom_point(alpha = 0.1) + geom_smooth(se=F)+xlab("alpha") +
  xlab(expression(alpha)) + ylab("Recall") +
  ggtitle(paste("Recall against", data_type)) +
  theme(
    strip.background = element_blank(),
    axis.title.x = element_text(size = 22),
    title = element_text(size = 16),
    strip.text = element_text(size = 16),
    legend.text = element_text(size = 15),
    axis.text = element_text(size = 12),
    legend.position = 'none'
  )
  return(precision/recall)
}

```

## ConnecTF

```{r, fig.width=10, fig.height=10}

load("results/100_permutations_lasso_validation_freq.rdata")

draw_precision_recall(val_conn, "ConnecTF")

# load(file = "results/100_permutations_lasso_validation_cluster_pos_freq.rdata")

draw_precision_recall(val_conn_pos, "ConnecTF, positive genes")
```

## CHIP Seq only

```{r, fig.width=10, fig.height=10, eval = F}
load("results/100_permutations_lasso_validation_chip.rdata")
draw_precision_recall(val_chip, "CHIP-Seq")

load(file = "results/100_permutations_lasso_validation_chip_cluster_pos.rdata")
# draw_precision_recall(val_chip, "CHIP-Seq, positive genes")
```


## TARGET only

```{r, fig.width=10, fig.height=10, eval = F}
load("results/100_permutations_lasso_validation_target.rdata")
draw_precision_recall(val_target, "TARGET")

load(file = "results/100_permutations_lasso_validation_target_cluster_pos.rdata")
# draw_precision_recall(val_target, "TARGET, positive genes")

```


## DAP Seq only

```{r, fig.width=10, fig.height=10}
load("results/100_permutations_lasso_validation_dap_mda_shuffle.rdata")
draw_precision_recall(val_dap, "DAP-Seq")
load("results/100_permutations_bRF_validation_dap.rdata")
draw_precision_recall(val_dap, "DAP-Seq")

# 
# load(file = "results/100_permutations_lasso_validation_dap_cluster_pos_freq.rdata")
# 
# draw_precision_recall(val_dap, "DAP-Seq, positive genes")
```

```{r, fig.width=10, fig.height=10}
draw_precision_recall(val_dap, "DAP-Seq, positive genes")
```

